### 2025-10-23 paper.qmd
more on python

---
title: "Nix for Polyglot, Reproducible Data Science Workflows"
format:
    jss-pdf:
        keep-tex: true
    jss-html:
        embed-resources: true
        toc: true
author:
  - name: Bruno Rodrigues
    affiliations:
      - name: Ministry of Research and Higher education, Luxembourg
        department: Department of Statistics
        address: 18, Montée de la Pétrusse
        city: Luxembourg
        country: Luxembourg
        postal-code: 2327
      - Journal of Statistical Software
    orcid: 0000-0002-3211-3689
    email: bruno@brodrigues.co
    url: https://www.brodrigues.co
abstract: |
  Reproducible analysis requires more than clean, well-documented code:
  researchers must also manage software dependencies, computational
  environments, and workflow execution. Existing tools like [Docker]{.pkg} and
  [renv]{.pkg} address parts of this challenge, but no single solution handles
  polyglot environments, system dependencies, and workflow orchestration
  together. I present two [R]{.proglang} packages that leverage [Nix]{.pkg} to
  provide both reproducible polyglot environments and declarative workflow
  orchestration. [rix]{.pkg} generates [Nix]{.pkg} expressions that define
  reproducible environments spanning [R]{.proglang}, [Python]{.proglang},
  [Julia]{.proglang}, and system dependencies. Building on these environments,
  [rixpress]{.pkg} (and its [Python]{.proglang} port, [ryxpress]{.pkg})
  orchestrates polyglot pipelines where each computational step runs in its own
  hermetically sealed environment, with automatic caching and dependency
  tracking. This approach enables researchers to develop analyses interactively
  while maintaining bit-for-bit reproducibility, supports collaboration across
  heterogeneous computational environments, and ensures that analyses remain
  executable years into the future.

keywords: [reproducibility, R, Nix]
keywords-formatted: [reproducibility, "[R]{.proglang}", "[Nix]{.pkg}"]
number-sections: true

bibliography: bibliography.bib
---

## Introduction: Reproducibility is also about software {#sec-intro}

@peng2011 introduced the concept of reproducibility as a *continuum*. At one end
lies the least reproducible state, where only a paper describing the study is
available. Reproducibility improves when authors share the original source code,
improves further when they include the underlying data, and reaches its highest
level when what Roger Peng called *linked and executable code and data* are
provided.

By *linked and executable code and data*, Peng referred to compiled source code
and runnable scripts. In this paper, I interpret this notion more broadly as the
*computational environment*: the complete set of software required to execute an
analysis. Here too, a continuum exists. At the minimal end, authors might only
name the main software used—say, the [R]{.proglang} programming language. More
careful authors might also specify the version of [R]{.proglang}, or list the
additional packages and their versions. Rarely, however, do authors specify the
operating system on which the analysis was performed, even though differences in
operating systems can lead to divergent results when using the same code and
software versions, as shown by @neupane2019. It is even less common for authors
to provide step-by-step installation instructions for the required software
stack, an omission often driven by institutional constraints. Journals, for
instance, can inadvertently work against reproducibility by imposing strict page
or word limits that leave no room for the necessary technical documentation,
effectively discouraging thoroughness in the name of brevity.

Even when such instructions are given, they often fail across different
platforms or versions of the same platform. This lack of portability not only
hinders reproducibility but also complicates everyday research workflows.
Researchers working across multiple machines must be able to recreate the same
environment consistently, and collaborators must share identical computational
setups to avoid inconsistencies.

Finally, once the execution environment is correctly configured, additional
clarity is needed on how to *run* the project itself. Which packages should be
loaded first? Which scripts should be executed, and in what order? Without clear
documentation or automated orchestration, these operational details become yet
another barrier to reproducibility.

This paper focuses specifically on two critical but often overlooked aspects of
reproducibility: *computational environment management* and *workflow
orchestration*. I present a comprehensive framework that addresses both
challenges using the [Nix]{.pkg} package manager, making it accessible to
researchers through two [R]{.proglang} packages: [rix]{.pkg} and
[rixpress]{.pkg}. Before introducing these packages, I survey existing tools and
their limitations to contextualize my contribution.

A range of tools now exist to help researchers approach the gold standard of
full reproducibility, or to consistently deploy the same development environment
across multiple machines. Let us first consider the most basic step in this
process: listing the software used. In [R]{.proglang}, the `sessionInfo()`
function provides a concise summary of the software environment, including the R
version, platform details, and all loaded packages. Its output can be saved to a
file and included as part of a study's reproducibility record. Below is an
example output from `sessionInfo()`:

::: {.content-hidden when-format="pdf"}
```R
sessionInfo()
```
:::

::: {.content-hidden when-format="pdf"}
```R
R version 4.3.2 (2023-10-31)
Platform: aarch64-unknown-linux-gnu (64-bit)
Running under: Ubuntu 22.04.3 LTS

Matrix products: default
BLAS:   /usr/lib/aarch64-linux-gnu/openblas-pthread/libblas.so.3
LAPACK: /usr/lib/aarch64-linux-gnu/openblas-pthread/libopenblasp[...]

locale:
 LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
 LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
 ...

attached base packages:
 stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
 nnet_7.3-19  mgcv_1.9-0   nlme_3.1-163

loaded via a namespace (and not attached):
 compiler_4.3.2  Matrix_1.6-1.1  tools_4.3.2     splines_4.3.2
 grid_4.3.2      lattice_0.21-9
```
:::

```{=tex}
\begin{CodeInput}
R> sessionInfo()
\end{CodeInput}
\begin{CodeOutput}
R version 4.3.2 (2023-10-31)
Platform: aarch64-unknown-linux-gnu (64-bit)
Running under: Ubuntu 22.04.3 LTS

Matrix products: default
BLAS:   /usr/lib/aarch64-linux-gnu/openblas-pthread/libblas.so.3
LAPACK: /usr/lib/aarch64-linux-gnu/openblas-pthread/libopenblasp[...]

locale:
 LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
 LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
 ...

attached base packages:
 stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
 nnet_7.3-19  mgcv_1.9-0   nlme_3.1-163

loaded via a namespace (and not attached):
 compiler_4.3.2  Matrix_1.6-1.1  tools_4.3.2     splines_4.3.2
 grid_4.3.2      lattice_0.21-9
\end{CodeOutput}
```

When an author includes this information, others attempting to reproduce the
study (future readers, collaborators, or even the author at a later time) can
easily see which version of [R]{.proglang} and which packages (with their
versions) were used. However, reproducing the environment still requires
manually installing the correct package versions—a process that becomes
particularly difficult when packages depend on system-level libraries. For
example, the [sf]{.pkg} package requires [GDAL]{.pkg}, [GEOS]{.pkg}, and
[PROJ]{.pkg} system libraries. Installing and configuring these dependencies
varies substantially across operating systems and can be prohibitively difficult
on some platforms.

A more robust approach than simply listing package versions is to use the
[renv]{.pkg} package. [renv]{.pkg} captures the project's software state and
writes it to a `renv.lock` file, which includes the exact versions of
[R]{.proglang} and all required packages. This lockfile serves as a blueprint
for restoring the environment automatically, ensuring that others can recreate
the same setup with minimal effort. Below is an abbreviated example of an
`renv.lock` file:
```{=tex}
\begin{Code}
{
  "R": {
    "Version": "4.2.2",
    "Repositories": [
      {
        "Name": "CRAN",
        "URL": "https://packagemanager.rstudio.com/all/latest"
      }
    ]
  },
  "Packages": {
    "MASS": {
      "Package": "MASS",
      "Version": "7.3-58.1",
      "Source": "Repository",
      "Repository": "CRAN",
      "Hash": "762e1804143a332333c054759f89a706",
      "Requirements": []
    },
    "Matrix": {
      "Package": "Matrix",
      "Version": "1.5-1",
      "Source": "Repository",
      "Repository": "CRAN",
      "Hash": "539dc0c0c05636812f1080f473d2c177",
      "Requirements": ["lattice"]
    }
    ... additional packages omitted ...
  }
}
\end{Code}
```

This lockfile lists each package alongside its version and the repository from
which it was downloaded. Creating it requires only a single command:
`renv::init()`. Others can then reproduce the same package library by running
`renv::restore()`, which installs the exact package versions in an isolated,
project-specific library that does not interfere with the user's global R
library.

However, [renv]{.pkg} has important limitations. It does not restore the version
of [R]{.proglang} itself—installing the correct R version must be done
separately using tools like [rig]{.pkg} [@rlib2023]. More critically,
[renv]{.pkg} does not handle system-level dependencies. If [sf]{.pkg} requires
[GDAL]{.pkg} version 3.0 but the system has version 2.4 installed, [renv]{.pkg}
cannot resolve this conflict. Users must manually install system libraries, and
the process differs across operating systems. Further details are discussed in
the [renv]{.pkg} documentation^[See
\url{https://rstudio.github.io/renv/articles/renv.html}.].

Other packages provide similar functionality with different trade-offs.
[groundhog]{.pkg} by @simonsohn2023 makes it easy to install packages as they
existed on CRAN at a given date:

::: {.content-hidden when-format="pdf"}
```R
groundhog.library("
 library('purrr')
 library('ggplot2')",
 "2017-10-04",
 tolerate.R.version = "4.2.2")
```
:::

```{=tex}
\begin{CodeInput}
R> groundhog.library("
+   library('purrr')
+   library('ggplot2')",
+   "2017-10-04",
+   tolerate.R.version = "4.2.2")
\end{CodeInput}
```

[groundhog]{.pkg} places packages in project-specific libraries but does not
install [R]{.proglang} itself. Without the `tolerate.R.version` argument, it
will refuse to proceed if the R version does not match expectations.
[rang]{.pkg}, developed by @chan2023, offers similar date-based package
installation. The Posit Package Manager provides another approach through dated
CRAN snapshots, configured via the `.Rprofile` file:

::: {.content-hidden when-format="pdf"}
```R
options(repos =
  c(REPO_NAME =
    "https://packagemanager.posit.co/cran/__linux__/jammy/2023-06-30"
  )
)
```
:::

```{=tex}
\begin{CodeInput}
R> options(repos =
+    c(REPO_NAME =
+      "https://packagemanager.posit.co/cran/__linux__/jammy/2023-06-30"
+    )
+  )
\end{CodeInput}
```

This approach installs all packages from the specified snapshot date, but unless
users explicitly manage separate libraries for different projects, all projects
will share the same package versions.

These tools represent significant progress in reproducibility, but they share a
common limitation: none addresses system-level dependencies or supports polyglot
environments. A project using [R]{.proglang}, [Python]{.proglang}, and system
tools like [Quarto]{.pkg} or \LaTeX requires coordinating multiple package
managers ([renv]{.pkg} for R, virtual environments for Python, manual
installation for system tools), each with its own configuration and potential
for conflicts.

The most comprehensive approach to reproducibility is containerization with
[Docker]{.pkg}. [Docker]{.pkg} packages a *data product* together with all its
dependencies—including the operating system, system libraries, programming
languages, and packages—into a self-contained image. A statistical analysis can
be viewed as such a data product. The steps to create an image are specified in
a `Dockerfile`, which defines the structure and content of the [Docker]{.pkg}
image.

Once the image is built, the analysis can be executed inside a *container*,
which is a running instance of the image. [Docker]{.pkg} containers are typically run
non-interactively, allowing the complete computational environment to be
instantiated and executed with a single command.

The strength of [Docker]{.pkg} lies in its ability to bundle not only the
correct versions of [R]{.proglang} and its packages but also system-level
dependencies. Since [Docker]{.pkg} images are effectively minimal Linux systems,
they naturally include libraries required by packages like [sf]{.pkg}. This
ensures that future replicators have access to the same software environment,
including all necessary system components. Moreover, these images can be easily
shared via registries like Docker Hub, enabling seamless reproducibility across
machines and collaborators.

The Rocker project, introduced by @boettiger2017, provides a collection of
pre-built [Docker]{.pkg} images for the [R]{.proglang} community. These images
come with specific R versions and, in some cases, preinstalled packages, making
them convenient base images for building reproducible analysis environments
without starting from scratch.

Despite its strengths, [Docker]{.pkg} presents several challenges for
statistical computing workflows:

1. Poor interactive support: While [Docker]{.pkg} containers can be modified at runtime,
  changes are lost when the container stops unless explicitly saved. Running
  graphical applications like RStudio Desktop from containers requires complex
  X11 configuration that works reliably only on Linux. Instead, web-based IDEs
  (like RStudio Server) are used, but come with the inherent limitations of
  web-based solutions.

2. Steep learning curve: [Docker]{.pkg} images are minimal Linux systems, so
  writing reliable `Dockerfile`s requires familiarity with Linux system
  administration. To ensure reproducibility, `Dockerfile`s should reference
  specific image *digests* rather than *tags*, but this practice is rarely
  followed. Packages like [dockerfiler]{.pkg} [@dockerfiler] help automate
  `Dockerfile` creation but cannot eliminate the need for basic Linux knowledge.

3. Post-hoc reproducibility: A common workflow is to develop analyses
  interactively using standard R installations, then create a `Dockerfile`
  afterward to enable reproduction. While this achieves post-hoc
  reproducibility, it fails to address the challenge of deploying consistent
  environments across multiple machines *during* development, a critical need
  for collaborative projects.

4. Single-language focus: While [Docker]{.pkg} can support polyglot
  environments, orchestrating them requires manual coordination. There is no
  standard way to specify that one step should run in a Python environment while
  another runs in R.

5. Incomplete Reproducibility: Reproducibility should be understood as a
  spectrum rather than a binary property, a notion emphasised in
  "Reproducibility in Software Engineering" by @dellaieraMasterThesis2024.
  Docker illustrates this spectrum well: it provides strong run-time
  reproducibility—that is, consistency in space (where the same image behaves
  identically across different systems) but weak build-time reproducibility,
  which concerns consistency over time. As defined in the thesis,
  reproducibility at build time means being able to produce an identical build
  artefact regardless of when or where it is built. However, Docker inherently
  fails to guarantee this property. Its reliance on mutable base images and
  non-deterministic instructions such as `apt-get update` leads to temporal
  drift: rebuilding the same `Dockerfile` at different points in time can
  produce distinct outputs. In the experimental evaluation conducted in the
  thesis, even with pinned base images, checksums of the resulting [Docker]{.pkg}
  containers varied across builds and machines. Thus, while Docker offers
  partial reproducibility: strong at run time (space) and weak at build time
  (time), it cannot on its own ensure verifiable, deterministic software
  artifacts. Achieving true reproducibility requires pinning image digests,
  freezing dependency versions, and employing deterministic build systems such
  as Nix or Guix for complete control over both space and time [@Malka_2024].

Even with a perfectly reproducible environment, researchers face another
challenge: reliably executing the analysis workflow. Which scripts run first?
What are the dependencies between steps? Manual execution is error-prone and
poorly documented. Build automation tools like [Make]{.pkg} address this by
defining analyses as ordered, reproducible steps.

Within [R]{.proglang}, the [targets]{.pkg} package by @landau2021 provides a
modern, declarative approach to workflow management. [targets]{.pkg} tracks
dependencies between code and data, caches intermediate results, and only
recomputes steps affected by changes. This dramatically improves efficiency for
complex analyses.

However, [targets]{.pkg} operates within a single R session. While it can call
Python via [reticulate]{.pkg} or execute system commands, all steps share the
same computational environment. For truly polyglot pipelines (where different
steps may require incompatible dependencies or different language versions) this
limitation becomes problematic. One might run [targets]{.pkg} inside a
[Docker]{.pkg} container to ensure reproducibility, but this requires the entire
pipeline to use a single environment.

@mcdermott2021 provides an excellent example of achieving the gold standard of
reproducibility. The accompanying GitHub
repository^[https://github.com/grantmcdermott/skeptic-priors] demonstrates the
state of the art:

- Package versions recorded in an `renv.lock` file
- A `Makefile` automating the full analysis
- A `Dockerfile` providing a complete computational environment

However, achieving this required mastering multiple complex tools: [renv]{.pkg}
for package management, [Docker]{.pkg} for environment containerization, and
[Make]{.pkg} for workflow orchestration. Each tool has its own syntax, concepts,
and failure modes. The complexity compounds for polyglot projects; for instance,
a [Python]{.proglang} project would require learning virtual environments, and
coordinating [R]{.proglang} and [Python]{.proglang} in a single analysis
requires additional tooling.

What researchers need is a tool that:

1. Manages complete environments including programming languages, packages, and
   system dependencies;
2. Supports multiple languages natively, not through workarounds;
3. Works interactively without the complexity of containers;
4. Provides step-level isolation so different pipeline steps can use
  different environments;
5. Ensures bit-for-bit reproducibility through deterministic builds;
6. Remains simple enough for researchers without systems administration
  expertise.

The [Nix]{.pkg} package manager provides these capabilities. [Nix]{.pkg} ensures
reproducible software installation by deploying *component closures*—packages
bundled with all their dependencies and transitive dependencies
[@dolstra2004nix]. Think of installing a package like packing for a trip.
Traditional package managers are like packing just your clothes, assuming you'll
find a toothbrush and soap at your destination. [Nix]{.pkg} packs everything you
might need—the entire self-contained kit. This is what [Nix]{.pkg} calls a
*component closure*.

This creates self-contained software environments that work
identically across machines and over time. [Nix]{.pkg} can replace
[Docker]{.pkg} for environment isolation, [renv]{.pkg} for package management,
and even [Make]{.pkg} for workflow orchestration—all within a single, unified
framework.

However, [Nix]{.pkg} has a steep learning curve. It is a complex system with its
own purely functional programming language designed to declaratively define how
software is built and configured. While this functional approach ensures
reproducibility, it creates a significant barrier to adoption.

To make [Nix]{.pkg} accessible to researchers, I developed two [R]{.proglang}
packages:

- [rix]{.pkg} generates [Nix]{.pkg} expressions from intuitive R function calls,
eliminating the need to learn the [Nix]{.proglang} language. It handles
environment definition including [R]{.proglang}, [Python]{.proglang},
[Julia]{.proglang}, system tools, and their dependencies.

- [rixpress]{.pkg} orchestrates polyglot analytical pipelines using [Nix]{.pkg}
as the build engine. Each pipeline step runs in its own hermetically sealed
environment with automatic caching and dependency tracking. This enables true
polyglot workflows where [R]{.proglang}, [Python]{.proglang} and
[Julia]{.proglang} steps coexist with different dependencies.
[Python]{.proglang} user can use [ryxpress]{.pkg}, a port of [rixpress]{.pkg}.

Together, these packages provide what existing tools cannot: a single framework
for managing reproducible, polyglot computational environments and executing
complex analytical workflows with step-level isolation. While [Nix]{.pkg}
remains complex for advanced use cases, [rix]{.pkg} and [rixpress]{.pkg}
abstract this complexity for common research workflows, making deep
reproducibility accessible without requiring systems administration expertise.

The remainder of this paper proceeds as follows. Section 2 introduces the
[Nix]{.pkg} package manager and explains why its functional approach enables
reproducibility. Section 3 presents [rix]{.pkg} and demonstrates environment
definition for various use cases. Section 4 discusses the `rstats-on-nix` fork
of the official package repository and how it addresses practical limitations.
Section 5 introduces [rixpress]{.pkg} and demonstrates polyglot pipeline
orchestration through a complete example. Section 6 concludes with discussion of
appropriate use cases and future directions.

## The Nix package manager {#sec-nix}

[Nix]{.pkg} is a powerful, cross-platform package manager designed for
installing and building software in a fully reproducible and reliable manner.
Unlike traditional package managers, [Nix]{.pkg} emphasizes immutable
infrastructure and a functional approach, which significantly enhances the
consistency of computational environments. As of writing, the primary Nix
package collection, `nixpkgs`, provides access to over 120,000 packages,
including nearly all of CRAN and Bioconductor. According to Repology, a service
that tracks package repositories across distributions, `nixpkgs` is both the
largest and among the most up-to-date package repositories available (see Figure
@fig-repology).

![Package repository size and freshness according to Repology. `nixpkgs` leads
in both total package count and update recency, making it well-suited for
reproducible research environments.](repology.png){#fig-repology}

This extensive coverage allows users to install not only [R]{.proglang} itself
but also all its required packages and system-level dependencies for any given
project.

While [Nix]{.pkg} is the default package manager for the NixOS Linux
distribution, it can be installed as a standalone tool on other Linux
distributions, macOS, and effectively on Windows via the Windows Subsystem for
Linux (WSL2), treating these as equivalent platforms in practice.

A key advantage of using [Nix]{.pkg} to install [R]{.proglang} packages, as
opposed to the standard `install.packages()` function, lies in its comprehensive
dependency management. [Nix]{.pkg} ensures that all dependencies of each package
are installed, irrespective of whether they are other [R]{.proglang} packages or
underlying system libraries.

Consider, for example, the [sf]{.pkg} package, which is fundamental for spatial
data analysis in [R]{.proglang}. [sf]{.pkg} relies on several complex external
system libraries such as [GDAL]{.pkg}, [GEOS]{.pkg}, and
[PROJ]{.pkg}. Manually installing and configuring these libraries to be
compatible with [sf]{.pkg} can be challenging and platform-specific. With
[Nix]{.pkg}, however, the user merely needs to declare [sf]{.pkg} as a project
requirement. [Nix]{.pkg} then automatically installs and configures all
necessary system libraries as transitive dependencies. This seamless process is
possible because the maintainers of the [R]{.proglang} programming language for
[Nix]{.pkg} have declaratively specified these dependencies for [sf]{.pkg} in
`nixpkgs`, allowing the entire software stack to be provisioned effortlessly
from an end-user perspective.

This concept is central to [Nix]{.pkg}'s design, where packages are referred to
as *component closures*. As @dolstra2004nix explains:

> The idea is to always deploy component closures: if we deploy a component,
> then we must also deploy its dependencies, their dependencies, and so on. That
> is, we must always deploy a set of components that is closed under the
> ''depends on'' relation. Since closures are selfcontained, they are the units
> of complete software deployment. After all, if a set of components is not
> closed, it is not safe to deploy, since using them might cause other
> components to be referenced that are missing on the target system.

The core of [Nix]{.pkg}'s reproducibility lies in its functional package
management paradigm and its unique approach to builds. When installing software
with [Nix]{.pkg}, it evaluates an *expression* written in the [Nix]{.proglang}
language. These expressions define *derivations*, which are declarative
blueprints describing how to build a package. A derivation specifies all inputs:
source code, build commands, and crucially, all direct and transitive
dependencies. This declarative nature ensures that every build process is fully
specified and independent of the environment where it's executed. Furthermore,
all [Nix]{.pkg} expressions for official packages are hosted in the `nixpkgs`
GitHub repository. By referencing a specific commit of `nixpkgs` (a process
known as *pinning a revision*) users guarantee that all packages installed are
derived from the exact same set of build instructions. This means that if an
environment is built today using a pinned revision, it will produce the
identical set of software versions and configurations if rebuilt tomorrow, next
year, or on a different machine, as long as the `nixpkgs` repository remains
accessible.

Pinning is essential, but it is not the only mechanism through which [Nix]{.pkg}
achieves reproducibility. [Nix]{.pkg} adopts principles from functional
programming, conceptualizing software builds as *pure functions*. In
mathematics, a pure function is defined as one that always produces the same
output for a given input—for instance, if \( f(x) = x^2 \), then \( f(2) = 4 \)
invariably holds. Analogously, [Nix]{.pkg} ensures that, given identical inputs
(such as source code, dependencies, and build instructions), the resulting
output (the built package) remains identical, irrespective of the environment or
time of execution. In this sense, [Nix]{.pkg} functions as a *functional package
manager*, applying the deterministic properties of pure functions to software
construction.

This is achieved by disallowing side effects and ensuring that builds operate
within hermetic (isolated) environments, with no hidden dependencies on global
system state. While this functional approach greatly enhances reproducibility,
it can sometimes introduce complexity for package maintainers, especially for
software that needs to download external assets during installation (e.g.,
certain Bioconductor packages). For end-users, however, this complexity is
largely abstracted away. Additionally, [Nix]{.pkg} supports multiple versions or
*variants* of a package on the same system.

[Nix]{.pkg} installs all packages into a centralized, content-addressed
directory called the [Nix]{.pkg} store (typically `/nix/store`). Each package
gets a unique subdirectory whose name includes a hash of all its inputs: source
code, dependencies, build instructions. This means
`/nix/store/abc123-dplyr-1.0.0` and `/nix/store/xyz789-dplyr-1.1.0` can coexist
without conflicts.

This isolation ensures that projects always use their intended software
versions, allowing multiple versions of [R]{.proglang} to coexist and preventing
"dependency hell" and global environment pollution.

For a more in-depth technical discussion of [Nix]{.pkg}'s design principles, see
@dolstra2004nix.

By leveraging these principles, [Nix]{.pkg} can effectively replace and
integrate the functionalities of tools like [renv]{.pkg} and [Docker]{.pkg} for
[R]{.proglang} projects, or `requirements.txt` files and virtual environments
for [Python]{.proglang} projects. Furthermore, [Nix]{.pkg} excels at building
polyglot environments, which can seamlessly combine [R]{.proglang},
[Python]{.proglang}, [Julia]{.proglang}, a [LaTeX]{.proglang} distribution, or
any of the numerous other tools available in `nixpkgs`. This enables the
creation of a truly complete, project-specific, and deeply reproducible
environment that can be used interactively for development or non-interactively
for automated analysis. As long as the `nixpkgs` repository (or a suitable fork,
as discussed in Section @sec-fork) remains accessible, the environment can be
rebuilt reliably in the future.

[Nix]{.pkg} provides strong cross-platform support across major operating
systems. It runs natively on Linux (x86_64) with full functionality and optimal
integration. On macOS (Intel and Apple Silicon), however, reproducibility is
less reliable due to dependencies on impure system frameworks and Xcode
toolchains outside the Nix store. As a result, macOS users may need to update
project pins after system or Xcode upgrades to restore builds. On Windows,
[Nix]{.pkg} works well through the Windows Subsystem for Linux (WSL2), though
graphical applications may need extra setup. [Nix]{.pkg} environments on WSL can
be used interactively via VS Code or Positron for a smooth development
experience. It is also important to note that [Nix]{.pkg} can be installed
within a [Docker]{.pkg} image to deploy and reproduce environments reliably in
containerized workflows.

While [Nix]{.pkg} is a powerful system, it comes with notable challenges. The
steepest is its learning curve: [Nix]{.pkg} uses its own declarative language
(also called [Nix]{.proglang}), which can be difficult for those new to
functional or declarative paradigms. Complexity becomes most apparent when
writing custom derivations or troubleshooting intricate build issues. Build
times can also be long, especially for large environments, when pre-built
binaries aren’t available. Additionally, because [Nix]{.pkg} stores all package
outputs immutably in the Nix store, disk usage tends to be higher than with
traditional package managers. Finally, while `nixpkgs` is extensive, some new or
niche packages may be missing or fail to build across platforms.

To make these capabilities more approachable for [R]{.proglang} users seeking
reproducible, project-specific environments without learning the full
[Nix]{.proglang} language, I created the [rix]{.pkg} and [rixpress]{.pkg}
packages.

## Reproducible development environments with Nix {#sec-repro-nix}

As mentioned, [Nix]{.pkg} expressions are written in the [Nix]{.proglang}
programming language, which is purely functional. Here is a simple example that
creates a shell environment containing version 4.3.1 of [R]{.proglang}:

::: {.content-hidden when-format="pdf"}
```nix
let
  pkgs = import (fetchTarball
    "https://github.com/NixOS/nixpkgs/archive/976fa336.tar.gz"
  ) {};
  system_packages = builtins.attrValues {
    inherit (pkgs) R;
  };
in
  pkgs.mkShell {
    buildInputs = [ system_packages ];
    shellHook = "R --vanilla";
  }
```
:::

```{=tex}
\begin{CodeInput}
let
  pkgs = import (fetchTarball
    "https://github.com/NixOS/nixpkgs/archive/976fa336.tar.gz"
  ) {};
  system_packages = builtins.attrValues {
    inherit (pkgs) R;
  };
in
  pkgs.mkShell {
    buildInputs = [ system_packages ];
    shellHook = "R --vanilla";
  }
\end{CodeInput}
```

In this expression, the `let` keyword is used to define variables. The variable
`pkgs` imports the set of packages from the `nixpkgs` repository at the
specified commit `976fa336`. The variable `system_packages` lists the packages
to include in the environment; in this case, it is just the [R]{.proglang}
programming language, along with all its dependencies and their transitive
dependencies. The `mkShell` function then creates a development shell with the
specified packages. The `shellHook` is set to `"R --vanilla"`, meaning that
entering the shell automatically starts [R]{.proglang} in vanilla mode, ignoring
any startup options.

This expression can be saved in a file called `default.nix`. The environment can
then be built on a system with [Nix]{.pkg} installed using the `nix-build`
command.^[For installing [Nix]{.pkg}, I recommend the Determinate Systems
installer: \url{https://determinate.systems/posts/determinate-nix-installer}].
Once the build completes, the user can enter the interactive shell with
`nix-shell`, which executes following steps:

- Nix reads the expression and identifies what needs to be installed;
- It checks if these packages already exist in `/nix/store`;
- If not, it downloads or builds them (with all dependencies);
- It creates a new shell environment and updates the `$PATH` to include the
  right binaries from `/nix/store`.
- The user is now in an interactive [Nix]{.pkg} shell session.

This shell contains all the packages specified in `default.nix` and
can be used for development, similar to activating a virtual environment in the
[Python]{.proglang} ecosystem.

Writing [Nix]{.pkg} expressions can be challenging for users unfamiliar with the
[Nix]{.proglang} language. However, the ability to define a fully reproducible
development environment in a single text file and then rebuild it anywhere is
highly appealing. To lower the barrier to adoption of [Nix]{.pkg} for
reproducibility, I developed the [rix]{.pkg} package.

[rix]{.pkg} provides the `rix()` function, which simplifies generating
[Nix]{.pkg} expressions. It is available on CRAN and can be installed like any
other R package. Additionally, it can bootstrap an [R]{.proglang} development
environment on a system where [R]{.proglang} is not yet installed but
[Nix]{.pkg} is available. This can be done by running (inside of a terminal):

::: {.content-hidden when-format="pdf"}
```bash
nix-shell -I \
  nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
  R rPackages.rix
```
:::

```{=tex}
\begin{CodeInput}
$> nix-shell -I \
+    nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
+    R rPackages.rix
\end{CodeInput}
```

(the `-I` flag allows one to pass a specific revision of `nixpkgs`, ensuring
temporary shells are also reproducible).

This command opens a temporary [R]{.proglang} session with [rix]{.pkg}
available.^[`nix-shell -p` starts an interactive shell with the specified
packages.] From there, users can generate new [Nix]{.pkg} expressions for
building environments. For example, the following generates a `default.nix` file
that installs [R]{.proglang} 4.3.1 along with the [dplyr]{.pkg} and
[chronicler]{.pkg} packages:

::: {.content-hidden when-format="pdf"}
```r
library('rix')

rix(r_ver = "4.3.1",
  r_pkgs = c("dplyr", "chronicler"),
  project_path = ".",
  overwrite = TRUE)
```
:::

```{=tex}
\begin{CodeInput}
R> library('rix')

R> rix(r_ver = "4.3.1",
+    r_pkgs = c("dplyr", "chronicler"),
+    project_path = ".",
+    overwrite = TRUE)
\end{CodeInput}
```

[rix]{.pkg} can also handle more complex setups, and users can provide a date
instead of a specific [R]{.proglang} version:

::: {.content-hidden when-format="pdf"}
```r
rix(date = "2025-10-20",
  r_pkgs = c("rix", "dplyr", "chronicler", "AER@1.2-8"),
  system_pkgs = c("quarto", "git"),
  tex_pkgs = c(
        "amsmath",
        "framed",
        "fvextra",
        "environ",
        "fontawesome5",
        "orcidlink",
        "pdfcol",
        "tcolorbox",
        "tikzfill"
  ),
  git_pkgs = list(
      package_name = "fusen",
      repo_url = "https://github.com/ThinkR-open/fusen",
      commit = "60346860111be79fc2beb33c53e195f97504a667"
  ),
  ide = "positron",
  project_path = ".",
  overwrite = TRUE)
```
:::

```{=tex}
\begin{CodeInput}
R> rix(date = "2025-10-20",
+    r_pkgs = c("rix", "dplyr", "chronicler", "AER@1.2-8"),
+    system_pkgs = c("quarto", "git"),
+    tex_pkgs = c(
+          "amsmath",
+          "framed",
+          "fvextra",
+          "environ",
+          "fontawesome5",
+          "orcidlink",
+          "pdfcol",
+          "tcolorbox",
+          "tikzfill"
+    ),
+    git_pkgs = list(
+        package_name = "fusen",
+        repo_url = "https://github.com/ThinkR-open/fusen",
+        commit = "60346860111be79fc2beb33c53e195f97504a667"
+    ),
+    ide = "positron",
+    project_path = ".",
+    overwrite = TRUE)
\end{CodeInput}
```

This call to `rix()` generates a `default.nix` file for a development shell that
encapsulates a complete and reproducible research environment. It provides
[R]{.proglang} and its packages (including [AER]{.pkg} at version 1.2-8),
several TeXLive packages for \LaTeX document authoring, development versions of
[rix]{.pkg} and [fusen]{.pkg} pulled directly from GitHub at a specific commit,
and the Positron editor. The reproducibility of this environment is guaranteed
by pinning all components to a single point in time: the [R]{.proglang} packages
are resolved from the CRAN snapshot of October 20, 2025, while all other system
tools are fixed to the version of `nixpkgs` from that same date. Typing
`nix-shell` in a terminal within the folder that contains this `default.nix`
will drop the user into a new shell. It should be noted that it uses the
`nixpkgs` from that same date. It is also possible to configure IDEs to
dynamically load this new shell to provide the proper development tooling.

It is also possible to include [Python]{.proglang} and [Julia]{.proglang}
packages using the `py_conf` and `jl_conf` arguments respectively.
[Julia]{.proglang} packages work the same way as [R]{.proglang} packages: they
are pinned to their versions as of the specified date.

[Python]{.proglang} packages, however, require special handling. Unlike CRAN or
the Julia package registry, PyPI (the [Python]{.proglang} Package Index)
contains over 500,000 packages with highly variable quality—many lack proper
build specifications, have broken dependencies, or are abandoned. Because of
this heterogeneity, [Nix]{.pkg} maintainers cannot automatically package all of
PyPI. Instead, they manually curate and package individual [Python]{.proglang}
packages that meet quality standards. This means some [Python]{.proglang}
packages, or specific versions, may not be available through [Nix]{.pkg}.

For projects requiring [Python]{.proglang} packages not yet available in
[Nix]{.pkg}, [rix]{.pkg} supports integration with [uv]{.pkg}, a modern
[Python]{.proglang} package manager. This allows users to combine Nix's
system-level reproducibility (Python interpreter, system libraries) with
[uv]{.pkg}'s comprehensive package availability. While this hybrid approach
sacrifices some of Nix's deterministic build guarantees for Python packages, it
provides a pragmatic solution when the needed packages are unavailable in
`nixpkgs`.

Users can include [uv]{.pkg} as a system package:

::: {.content-hidden when-format="pdf"}
```r
rix(date = "2025-10-20",
  r_pkgs = c("rix", "dplyr", "chronicler"),
  system_pkgs = c("uv"),
  project_path = ".",
  overwrite = TRUE)
```
:::
```{=tex}
\begin{CodeInput}
R> rix(date = "2025-10-20",
+    r_pkgs = c("rix", "dplyr", "chronicler"),
+    system_pkgs = c("uv"),
+    project_path = ".",
+    overwrite = TRUE)
\end{CodeInput}
```

This hybrid approach offers complementary strengths: [Nix]{.pkg} manages the
[Python]{.proglang} interpreter, system libraries (e.g., [GDAL]{.pkg},
[HDF5]{.pkg}), and [R]{.proglang} packages with full determinism, while
[uv]{.pkg} handles [Python]{.proglang} packages through its own lockfile
mechanism (`uv.lock`). Within the Nix environment, users run standard [uv]{.pkg}
commands (e.g., `uv pip install pandas==2.0.0`) to install [Python]{.proglang}
packages. The [uv]{.pkg} lockfile records exact package versions and their
hashes, providing reproducibility for the [Python]{.proglang} package layer.

While this approach doesn't achieve Nix's level of build-time determinism for
[Python]{.proglang} packages ([uv]{.pkg} still downloads wheels or builds
packages at install time rather than using pre-built, content-addressed
artifacts) it provides a pragmatic balance: comprehensive [Python]{.proglang}
package availability with reasonable reproducibility guarantees. For polyglot
projects, this means [R]{.proglang} and system-level dependencies remain fully
reproducible through [Nix]{.pkg}, while [Python]{.proglang} packages gain the
reproducibility that [uv]{.pkg} can provide.

[rix]{.pkg} can generate [Nix]{.pkg} expressions even if [Nix]{.pkg} is not
installed on the system. This is useful for continuous integration and
continuous deployment (CI/CD) workflows on platforms such as GitHub Actions. For
instance, the repository containing the source code for this
article^[https://github.com/b-rodrigues/rix_paper] uses GitHub Actions to
compile the paper. Each time a push is made to the master branch, a runner
installs [Nix]{.pkg}, generates the environment from the hosted `default.nix`
file, and compiles the paper using [Quarto]{.pkg} within the reproducible
environment. This ensure that *exactly* the same environment is used on the
author's computer and on the CI/CD without any additional, platform-specific,
configuration.

Instead of first entering a [Nix]{.pkg} shell, it is also possible to run a
program directly from the environment:

::: {.content-hidden when-format="pdf"}
```bash
cd /path/to/project/ && nix-shell default.nix --run "Rscript analysis.R"
```
:::

```{=tex}
\begin{CodeInput}
cd /path/to/project/ && nix-shell default.nix --run "Rscript analysis.R"
\end{CodeInput}
```

This command runs `Rscript` and executes the `analysis.R` script, which in this
example should be located in the same directory as `default.nix`.

## The rstats-on-nix fork of nixpkgs {#sec-fork}

As explained earlier, [Nix]{.pkg} uses expressions from the `nixpkgs` GitHub
repository to build software. However, when generating expressions with
[rix]{.pkg}, the fork `rstats-on-nix/nixpkgs` is used instead.

Using a fork offers several advantages. First, it provides flexibility that the
official `nixpkgs` repository cannot always accommodate. [Nix]{.pkg} is
primarily the package manager for the NixOS Linux distribution, and governance
and technical choices made upstream can limit what [rix]{.pkg} aims to provide.

For instance, while [Nix]{.pkg} can theoretically support multiple versions (or
*variants*) of the same package, in practice maintainers cannot provide several
variants for all [R]{.proglang} packages, given the size of the ecosystem (over
20,000 CRAN and Bioconductor packages). This makes it difficult to install a
specific version of an [R]{.proglang} package not included in a particular
`nixpkgs` commit. With [rix]{.pkg}, users can install a specific package version
from source, e.g.:

::: {.content-hidden when-format="pdf"}
```r
rix(..., r_pkgs = "dplyr@1.0.7", ...)
```
:::

```{=tex}
\begin{CodeInput}
R> rix(..., r_pkgs = "dplyr@1.0.7", ...)
\end{CodeInput}
```

However, installing from source might fail, especially if the package needs to
be compiled.

Additionally, updating the full [R]{.proglang} package set on [Nix]{.pkg} daily
is impractical. While CRAN and Bioconductor update daily, the [R]{.proglang}
packages in `nixpkgs` are only updated with new [R]{.proglang} releases. This
limitation is due to [Nix]{.pkg}’s governance as a Linux distribution package
manager.

The `rstats-on-nix` fork allows me to circumvent these limitations. For example,
it provides daily snapshots of CRAN. Each day, the [R]{.proglang} package set is
updated and committed to a dated branch using GitHub Actions. Users can select a
specific date with:

::: {.content-hidden when-format="pdf"}
```r
rix(date = "2024-12-14", ...)
```
:::

```{=tex}
\begin{CodeInput}
R> rix(date = "2024-12-14", ...)
\end{CodeInput}
```

I strive to provide an available date per week: each Monday, a GitHub Action
tests popular packages on Linux and macOS, and only if all tests succeed is the
date added to the list of available dates in [rix]{.pkg}. This ensures users can
reliably install packages, and allows me to backport fixes if needed. For
example, when RStudio was temporarily broken due to a dependency issue
(`boost`), a pull request was submitted to the official `nixpkgs` repository. I
backported the fix to the `rstats-on-nix` fork, making RStudio available to
users of [rix]{.pkg} earlier than upstream, as merging PRs in the official
repository can take some time.

I have backported fixes to the `rstats-on-nix` `nixpkgs` fork as far back as
March 2019. The process involves checking out a `nixpkgs` commit on the selected
date, updating the [R]{.proglang} package set using Posit CRAN and Bioconductor
snapshots, backporting fixes, and ensuring popular packages work on both
x86-linux (including WSL2) and aarch64-darwin (Apple Silicon). These changes are
committed to a dated branch in `rstats-on-nix/nixpkgs`. Users can see all
available dates with `rix::available_dates()`.

A drawback of forking `nixpkgs` is that backported packages are not included
upstream and thus are not prebuilt by Nix’s CI platform, Hydra. Users may need
to build many packages from source, which can be time-consuming. To mitigate
this, I provide a binary cache sponsored by [Cachix](https://www.cachix.org/),
complementing the public Nix cache. Instructions for using Cachix are in
[rix]{.pkg}’s documentation. Using the cache significantly speeds up
installations, as prebuilt packages are downloaded rather than compiled.

The fork also allows me to catch issues (such as packages’ builds breaking) early
on, and prepare fixes that can then be contributed upstream.

## Orchestrating the workflow with rixpress {#sec-rixpress}

Defining a reproducible environment with [rix]{.pkg} addresses the first major
challenge of reproducibility. The second challenge is reliably and efficiently
executing the analysis workflow within that environment, which is the role of
the [rixpress]{.pkg} package (or [ryxpress]{.pkg} for [Python]{.proglang}
users).

As mentioned in the introduction, a build automation tool like [targets]{.pkg}
is invaluable for managing complex analyses. It tracks dependencies between code
and data, caches results, and only recomputes steps that have changed. One can
run a [targets]{.pkg} pipeline inside a Nix environment to make it reproducible.
However, this approach has limitations: the entire pipeline must run in a single
environment, and orchestrating steps across different languages (e.g.,
[R]{.proglang} and [Python]{.proglang}) requires manual handling via packages
like [reticulate]{.pkg}.

[rixpress]{.pkg} overcomes these limitations by using Nix not just as a package
manager, but as the build automation engine itself. In a [rixpress]{.pkg}
pipeline, each step is defined as a Nix derivation, providing two key benefits:

  1. True Polyglot Pipelines: Each step can have its own Nix environment. A
     Python step can run in a pure Python environment, an [R]{.proglang} step in
     an [R]{.proglang} environment, and a Quarto rendering step in yet another,
     all within the same pipeline.
  2. Deep Reproducibility: Each step is a hermetically sealed Nix derivation
     whose output is cached in the Nix store based on the hash of all its
     inputs. All artifacts are direct children of the computational environment
     (because the computational environment is actually a dependency of the
     artifacts), ensuring they are rebuilt if the environment changes, keeping
     environment and outputs always in sync.

To illustrate these capabilities with a realistic research task, I present a
polyglot pipeline that simulates a Real Business Cycle (RBC) model in
[Julia]{.proglang} [@bezanson2017julia], uses the resulting data to train an
[XGBoost]{.pkg} forecasting model in [Python]{.proglang}, visualizes the results
in [R]{.proglang} with [ggplot2]{.pkg}, and finally compiles a [Quarto]{.pkg}
report. The code for this example can be found in the [following
repository](https://github.com/b-rodrigues/rixpress_demos/tree/master/rbc)^[https://github.com/b-rodrigues/rixpress_demos/tree/master/rbc],
with additional details provided in the Appendix of this manuscript.

For new projects, [rixpress]{.pkg} provides `rxp_init()`, which generates the
necessary boilerplate structure including a pipeline definition file and
template helper scripts. Users then define their pipeline in the script
`gen-pipeline.R` using functions inspired by [targets]{.pkg}, and the
computational environment the pipeline will be executed in the `gen-env.R`
script. If functions are needed for pipeline steps, these are to be specified, in
language-specific helper scripts: `functions.jl` for [Julia]{.proglang},
`functions.py` for [Python]{.proglang}, and `functions.R` for [R]{.proglang}.
This separation keeps the pipeline definition clean and readable while
encapsulating the scientific logic in appropriate language-native files.
Importantly, these helper scripts are standard language files—they contain no
[rixpress]{.pkg}-specific code and can be reused in other projects or executed
independently outside of [Nix]{.pkg}.

The following example demonstrates how [rixpress]{.pkg} orchestrates a granular,
multi-step data science workflow that crosses language boundaries. (Complete
setup instructions and a detailed walkthrough are provided in Appendix Section
@sec-polyglot-example.)

::: {.content-hidden when-format="pdf"}
```r
library('rixpress')

pipeline_steps <- list(
  # STEP 0: Define RBC Model Parameters in Julia
  rxp_jl(alpha, 0.3),
  rxp_jl(beta, 1 / 1.01),
  # ... (other parameters omitted for brevity) ...

  # STEP 1: Julia - Simulate the RBC model
  rxp_jl(
    name = simulated_rbc_data,
    expr = "simulate_rbc_model(alpha, beta, delta, rho, sigma, sigma_z)",
    user_functions = "functions/functions.jl",
    encoder = "arrow_write"
  ),

  # STEP 2.1: Python - Prepare features
  rxp_py(
    name = processed_data,
    expr = "prepare_features(simulated_rbc_data)",
    user_functions = "functions/functions.py",
    decoder = "pyarrow.feather.read_feather"
  ),

  # STEP 2.2: Python - Split data (X_train, y_train, etc.)
  rxp_py(name = X_train, expr = "get_X_train(processed_data)", ...),
  # ... (other data splits omitted for brevity) ...

  # STEP 2.3: Python - Train the XGBoost model
  rxp_py(
    name = trained_model,
    expr = "train_model(X_train, y_train)",
    user_functions = "functions/functions.py"
  ),

  # STEP 2.4: Python - Make predictions and format results
  # ... (prediction and formatting steps omitted for brevity) ...
  rxp_py(
    name = final_predictions_df,
    expr = "format_results(y_test, model_predictions)",
    user_functions = "functions/functions.py",
    encoder = "save_arrow"
  ),

  # STEP 3: R - Visualize the predictions
  rxp_r(
    name = output_plot,
    expr = plot_predictions(final_predictions_df),
    user_functions = "functions/functions.R",
    decoder = arrow::read_feather
  ),

  # STEP 4: Quarto - Compile the final report
  rxp_qmd(
    name = final_report,
    qmd_file = "readme.qmd"
  )
)

# Generate the 'pipeline.nix' file from the R list
rxp_populate(pipeline_steps, build = TRUE)
```
:::

```{=tex}
\begin{CodeInput}
R> library('rixpress')

R> pipeline_steps <- list(
+  # STEP 0: Define RBC Model Parameters in Julia
+  rxp_jl(alpha, 0.3),
+  rxp_jl(beta, 1 / 1.01),
+  # ... (other parameters omitted for brevity) ...
+
+  # STEP 1: Julia - Simulate the RBC model
+  rxp_jl(
+    name = simulated_rbc_data,
+    expr = "simulate_rbc_model(alpha, beta, delta, rho, sigma, sigma_z)",
+    user_functions = "functions/functions.jl",
+    encoder = "arrow_write"
+  ),
+
+  # STEP 2.1: Python - Prepare features
+  rxp_py(
+    name = processed_data,
+    expr = "prepare_features(simulated_rbc_data)",
+    user_functions = "functions/functions.py",
+    decoder = "pyarrow.feather.read_feather"
+  ),
+
+  # STEP 2.2: Python - Split data (X_train, y_train, etc.)
+  rxp_py(name = X_train, expr = "get_X_train(processed_data)", ...),
+  # ... (other data splits omitted for brevity) ...
+
+  # STEP 2.3: Python - Train the XGBoost model
+  rxp_py(
+    name = trained_model,
+    expr = "train_model(X_train, y_train)",
+    user_functions = "functions/functions.py"
+  ),
+
+  # STEP 2.4: Python - Make predictions and format results
+  # ... (prediction and formatting steps omitted for brevity) ...
+  rxp_py(
+    name = final_predictions_df,
+    expr = "format_results(y_test, model_predictions)",
+    user_functions = "functions/functions.py",
+    encoder = "save_arrow"
+  ),
+
+  # STEP 3: R - Visualize the predictions
+  rxp_r(
+    name = output_plot,
+    expr = plot_predictions(final_predictions_df),
+    user_functions = "functions/functions.R",
+    decoder = arrow::read_feather
+  ),
+
+  # STEP 4: Quarto - Compile the final report
+  rxp_qmd(
+    name = final_report,
+    qmd_file = "readme.qmd"
+  )
+)

# Generate the 'pipeline.nix' file from the R list
R> rxp_populate(pipeline_steps, build = TRUE)
\end{CodeInput}
```

Computation steps, referred to as *derivations*, are defined using the functions
`rxp_r()`, `rxp_jl()`, and `rxp_py()`, corresponding to [R]{.proglang},
[Julia]{.proglang}, and [Python]{.proglang} environments, respectively. Variants
of these functions suffixed with `_file()` designate *import derivations*, which
introduce external data into the pipeline. It is important to note that any data
imported in this way is stored within the Nix store, ensuring reproducibility
but also persisting the data in the build environment. The function `rxp_qmd()`
is used to compile a [Quarto]{.pkg} document (there is also `rxp_rmd()` which
serves the same purpose for [R Markdown]{.pkg} documents).

The `rxp_populate()` function translates this [R]{.proglang} list into a
`pipeline.nix` file, which declaratively defines the entire workflow. Data flows
from derivation to derivation by being serialized into the efficient and
language-agnostic Arrow format using the pair of `encoder/decoder` functions.
Other universal formats, such as `csv` or `json` could have been used.

As a sidenote: [Python]{.proglang} users who wish to use [ryxpress]{.pkg} define
pipelines using the same [R]{.proglang}-based DSL shown above. This design
choice keeps the pipeline definition language consistent across both
[R]{.proglang} and [Python]{.proglang} ecosystems. [ryxpress]{.pkg} calls
[R]{.proglang} and [rixpress]{.pkg} under the hood to generate and build the
`pipeline.nix` file. To use [ryxpress]{.pkg}, one should simply add this
package to the `py_conf` argument of `rix()`.

![Graphical representation of the polyglot pipeline. Purple nodes are
[Julia]{.proglang}, yellow nodes are [Python]{.pkg}, light blue nodes are
[R]{.proglang}, and dark blue nodes are Quarto derivations.](dag.png){#fig-dag}

The package can then generate a visual representation of the pipeline's directed
acyclic graph by running `rxp_dag()`, as can be seen in @fig-dag.

To execute the pipeline, one can either set `build = TRUE` in `rxp_populate()`
or call `rxp_make()` separately. [Nix]{.pkg} executes each step in order,
building dependencies as needed. Outputs are cached, so subsequent runs only
recompute steps with changed inputs or code. This provides the efficiency of
[targets]{.pkg} with polyglot support and bit-for-bit reproducibility. Artifacts
can be inspected interactively in [R]{.proglang} using
`rxp_read("artifact_name")` or `rxp_load("artifact_name")`.

[rixpress]{.pkg} also includes several additional features not covered here for
brevity. As mentioned previously, it is also possible to configure popular IDEs
to work interactively and seamlessly with both [rix]{.pkg} and [rixpress]{.pkg},
enabling a smooth, reproducible workflow from within the development
environment. Detailed setup instructions are provided in the vignettes of both
packages.

## Conclusion {#sec-conclusion}

Many tools exist to improve reproducibility, but [Nix]{.pkg} stands out because
it deploys complete software environments *closed under the "depends on"
relation*: it installs not only a package, but all its dependencies and their
dependencies. This makes [Nix]{.pkg} particularly powerful for reproducible
research.

However, solving such a complex problem makes [Nix]{.pkg} a complex tool. With
[rix]{.pkg}, I aim to make [Nix]{.pkg} more accessible to [R]{.proglang} users
by providing a familiar interface and workflow. By building reproducible
development shells with [Nix]{.pkg}, researchers can accommodate a wide range of
use cases: running scripts and pipelines, developing interactive [shiny]{.pkg}
applications, or serving [plumber]{.pkg} APIs.

Furthermore, [rixpress]{.pkg} extends this reproducibility to entire analysis
pipelines. By leveraging Nix as a build automation engine, [rixpress]{.pkg}
allows polyglot workflows where each step runs in its own hermetically sealed
environment. This ensures deep reproducibility, efficient caching, and seamless
orchestration of polyglot analyses. Most importantly, it provides native
polyglot support without the manual coordination required by containerization
approaches, making complex multi-language workflows accessible to researchers
without systems administration expertise.

While [rix]{.pkg} and [rixpress]{.pkg} (and [ryxpress]{.pkg}) provide
significant steps forward, it is important to acknowledge current limitations.
[rixpress]{.pkg} (and [ryxpress]{.pkg}) in particular requires further
development, especially to enhance the ease of debugging complex pipelines. Its
visualization capabilities are also still fairly limited; for instance, they do
not currently indicate outdated derivations when a user modifies code that would
necessitate a rebuild. These areas represent important avenues for future work
to further mature the framework and improve the user experience.

## Acknowledgments {.unnumbered}

I thank the rOpenSci reviewers and contributors who provided valuable feedback
on the development of [rix]{.pkg} and [rixpress]{.pkg}. In particular, I am
grateful to David Watkins and Jacob Wujiciak-Jens for their reviews of
[rix]{.pkg}, and to William Landau and Anthony Martinez for their reviews of
[rixpress]{.pkg}. I also acknowledge the contributions of Philipp Baumann,
[rix]{.pkg}’s co-author, Richard J. Acton, Jordi Rosell, Elio Campitelli, László
Kupcsik, and Michael Heming for [rix]{.pkg}. Their expertise and feedback
greatly improved the quality and usability of these packages. I would also like
to thank Pol Dellaiera and Edvin Syk for providing feedback on the manuscript.

## References {.unnumbered}

:::{#refs}

:::

## Appendix {#sec-appendix}

### Reproducing this paper

The source code of this paper is hosted on GitHub and can be found
at following
[link](https://github.com/b-rodrigues/rix_paper)^[https://github.com/b-rodrigues/rix_paper].
The paper can be easily compiled by running the following command:

::: {.content-hidden when-format="pdf"}
```bash
nix-shell --run "quarto render paper.qmd --to jss-pdf"
```
:::

```{=tex}
\begin{CodeInput}
$> nix-shell --run "quarto render paper.qmd --to jss-pdf"
\end{CodeInput}
```

The `default.nix` file defines the exact computational environment required to
compile the manuscript, ensuring full reproducibility of the build process. To
guarantee the reproducibility of the manuscript, it is automatically
recompiled via GitHub Actions upon each commit, using the same Nix environment.
An HTML version is additionally deployed to GitHub Pages, providing an
accessible format for viewing on smaller devices at this
[link](https://b-rodrigues.github.io/rix_paper/)^[[https://b-rodrigues.github.io/rix_paper/](https://b-rodrigues.github.io/rix_paper/)].
The PDF version of the manuscript can be found at
[link](https://b-rodrigues.github.io/rix_paper/paper.pdf)^[[https://b-rodrigues.github.io/rix_paper/paper.pdf](https://b-rodrigues.github.io/rix_paper/paper.pdf)].

### A Complete Polyglot Example with rixpress

This appendix provides a conceptual walk-through of the polyglot pipeline
example discussed in Section 5. The pipeline simulates a Real Business Cycle
(RBC) model in [Julia]{.proglang}, trains an [XGBoost]{.pkg} model in
[Python]{.proglang}, visualizes the results in [R]{.proglang}, and compiles a
final report with [Quarto]{.pkg}.

The complete, runnable source code for this example is available in the paper's
GitHub repository. A shell script, `run_polyglot_example.sh`, is provided to
automate the entire process from environment creation to final output, as
described at the end of this
section^[\url{https://github.com/b-rodrigues/rix_paper/blob/master/polyglot-example/run_polyglot_example.sh}].

#### Project Structure and Components

The project is organized into a set of scripts, each with a distinct
responsibility. The core logic is separated from the environment definition and
pipeline orchestration.

*   `functions/`: This directory contains the helper scripts with the core analytical code for each language.
    *   `functions.jl`: The [Julia]{.proglang} script for the economic simulation.
    *   `functions.py`: The [Python]{.proglang} script for the machine learning workflow.
    *   `functions.R`: The [R]{.proglang} script for data visualization.
*   `gen-env.R`: An [R]{.proglang} script that defines the reproducible computational environment.
*   `gen-pipeline.R`: The main [R]{.proglang} script that defines and orchestrates the entire polyglot pipeline.

#### Step 1: The Environment Definition

The foundation of the project is the reproducible environment, defined in
`gen-env.R`. This script uses the `rix()` function to programmatically generate
a `default.nix` file. This file serves as a complete blueprint for the
computational environment, specifying:

*   The exact versions of [R]{.proglang}, [Python]{.proglang}, and
    [Julia]{.proglang}.
*   A list of required packages for each language (e.g., [ggplot2]{.pkg} for
  [R]{.proglang}, [xgboost]{.pkg} for [Python]{.proglang}, [DataFrames]{.pkg}
  for [Julia]{.proglang}).
*   System-level dependencies like [Quarto]{.pkg}.

Crucially, the entire environment is pinned to a specific date (`2025-10-14`),
ensuring that anyone who builds the environment, now or in the future, will get
the exact same software versions, guaranteeing reproducibility.

By dropping into a temporary shell using the following command:

::: {.content-hidden when-format="pdf"}
```bash
nix-shell -I \
  nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
  R rPackages.rix
```
:::

```{=tex}
\begin{CodeInput}
$> nix-shell -I \
+  nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
+  R rPackages.rix
\end{CodeInput}
```

it is possible to generate the `default.nix` by sourcing `gen-env.R`. Then, one
leaves this temporary shell, and builds the environment using the command
`nix-shell`, which also drop the user into the development shell.

#### Step 2: The Core Analytical Logic

The scientific logic for each stage of the pipeline is encapsulated in the
separate helper scripts within the `functions/` directory.

*   The RBC Model Simulation (`functions.jl`): This [Julia]{.proglang} script
  contains a single pure function that implements the state-space solution to
  the RBC model, based on @depaoli2009rbc. It takes the model's economic
  parameters as inputs and returns the simulated time-series data as a
  [DataFrame]{.pkg}.
*   The [XGBoost]{.pkg} Forecasting Model (`functions.py`): This
  [Python]{.proglang} script handles the complete machine learning workflow
  through a series of modular functions. Its responsibilities include feature
  engineering (creating lagged variables), splitting the data into training and
  testing sets, training the [XGBoost]{.pkg} model, and generating predictions.
*   The Visualization (`functions.R`): This [R]{.proglang} script contains a
  function that uses [ggplot2]{.pkg} to create the final visualization. It is
  designed to take the data frame of actual and predicted values produced by the
  Python script and generate a plot comparing the two series.

#### Step 3: Orchestrating the Pipeline

The entire workflow is defined and orchestrated by `gen-pipeline.R`. This script
acts as the master plan, using functions from the [rixpress]{.pkg} package to
define each computational step as a *derivation*. It declaratively outlines the
dependencies between steps:

1.  It begins by defining the RBC model parameters in [Julia]{.proglang}.
2.  It specifies that the RBC simulation in `functions.jl` depends on these
    parameters.
3.  It then defines the series of [Python]{.proglang} steps (feature
  preparation, training, prediction), making each one dependent on the output of
  the previous one. The first [Python]{.proglang} step is explicitly made
  dependent on the output from the [Julia]{.proglang} simulation.
  [rixpress]{.pkg} handles the passing of data between languages, in this case
  using the [Arrow]{.pkg} file format for efficiency.
4.  Finally, it defines the **R visualization** step, which depends on the final
  predictions from the [Python]{.proglang} model, and a **Quarto rendering**
  step that depends on the generated plot.

When this script is run in the development shell (by executing
`source("gen-pipeline.R")`), [rixpress]{.pkg} translates the declared pipeline
into a master [Nix]{.proglang} expression that [Nix]{.pkg} can execute,
automatically handling the caching of results and re-running only the necessary
steps if a piece of code or data changes.

To build the pipeline from an interactive [Python]{.proglang} session, one would
execute the following lines:

::: {.content-hidden when-format="pdf"}
```python
from ryxpress import rxp_make
rxp_make()
```
:::

```{=tex}
\begin{CodeInput}
Python> from ryxpress import rxp_make
Python> rxp_make()
\end{CodeInput}
```

#### Running the Project

This entire example can be executed by running the `run_polyglot_example.sh`
script available in the root of the paper's repository.
This script automates the full process:
1.  It first executes `gen-env.R` inside a temporary [Nix]{.pkg} shell to build
    the `default.nix` file.
2.  It then executes `gen-pipeline.R` inside the newly defined environment. This
  triggers [Nix]{.pkg} to run the entire polyglot pipeline in the correct order.

Upon completion, the script will have generated all intermediate artifacts and
the final `readme.html` report found in the `pipeline-output` folder containing
the visualization.