### 2025-10-26 docker-polyglot-example/rbc/readme.qmd
docker example

---
title: "RBC Model Analysis Report"
format: 
  html:
    embed-resources: true
    toc: true
---

## Model Performance

The plot below compares the actual output from the RBC model simulation against the predictions generated by the XGBoost model.

```{r}
#| echo: false
#| label: rbc-plot
#| fig-cap: "Comparison of Actual RBC Model Output vs. XGBoost Predictions"

# This chunk now simply includes the plot that was generated by the 
# 03_visualize_results.R script in the pipeline.
knitr::include_graphics("plots/output_plot.png")
```