### 2025-10-27 paper.qmd
typo

---
title: "Nix for Polyglot, Reproducible Data Science Workflows"
format:
    jss-pdf:
        keep-tex: true
    jss-html:
        embed-resources: true
        toc: true
author:
  - name: Bruno Rodrigues
    affiliations:
      - name: Ministry of Research and Higher education, Luxembourg
        department: Department of Statistics
        address: 18, Montée de la Pétrusse
        city: Luxembourg
        country: Luxembourg
        postal-code: 2327
      - Journal of Statistical Software
    orcid: 0000-0002-3211-3689
    email: bruno@brodrigues.co
    url: https://www.brodrigues.co
  - name: Philipp Baumann
    affiliations:
      - name: Alliance SwissPass, Switzerland
        department: Data and Analytics
        address: 7, Länggassstrasse
        city: Bern
        country: Switzerland
        postal-code: 3012
    orcid: 0000-0002-3194-8975
    email: baumann-philipp@protonmail.com
abstract: |
  Reproducible analysis requires more than clean, well-documented code; it
  demands the precise management of software dependencies, computational
  environments, and workflow execution. While researchers can combine tools like
  [Docker]{.pkg}, [renv]{.pkg} for [R]{.proglang} packages (or [uv]{.pkg} for
  [Python]{.proglang} packages) and [Make]{.pkg} to address these needs, the
  resulting toolchain is often complex and fragile, requiring the coordination
  of multiple disparate systems. This paper introduces a unified framework built
  on the [Nix]{.pkg} package manager, made accessible through two [R]{.proglang}
  packages. The first, [rix]{.pkg}, generates declarative [Nix]{.pkg}
  expressions to define version-pinned, reproducible environments that encompass
  [R]{.proglang}, [Python]{.proglang}, [Julia]{.proglang}, and system-level
  dependencies. The second, [rixpress]{.pkg} (and its [Python]{.proglang} port,
  [ryxpress]{.pkg}), leverages these environments to orchestrate polyglot
  pipelines where each computational step runs in a hermetically sealed,
  language-specific environment with automatic caching. By treating the entire
  computational environment as code, this integrated approach ensures a high
  degree of reproducibility, supports collaboration across heterogeneous
  systems, and helps guarantee that analyses remain executable long into the
  future.

keywords: [reproducibility, R, Nix]
keywords-formatted: [reproducibility, "[R]{.proglang}", "[Python]{.proglang}", "[Julia]{.proglang}", "[Nix]{.pkg}"]
number-sections: true

bibliography: bibliography.bib
---

## Introduction: Reproducibility is also about software {#sec-intro}

@peng2011 introduced the concept of reproducibility as a *continuum*. At one end
lies the least reproducible state, where only a paper describing the study is
available. Reproducibility improves when authors share the original source code,
improves further when they include the underlying data, and reaches its highest
level when what Roger Peng called *linked and executable code and data* are
provided.

By *linked and executable code and data*, Peng referred to compiled source code
and runnable scripts. In this paper, we interpret this notion more broadly as the
*computational environment*: the complete set of software required to execute an
analysis. Here too, a continuum exists. At the minimal end, authors might only
name the main software used—say, the [R]{.proglang} programming language
[@r-core-team]. More careful authors might also specify the version of
[R]{.proglang}, or list the additional packages and their versions. Rarely,
however, do authors specify the operating system on which the analysis was
performed, even though differences in operating systems can lead to divergent
results when using the same code and software versions, as shown by
@neupane2019. It is even less common for authors to provide step-by-step
installation instructions for the required software stack, an omission often
driven by institutional constraints. Journals, for instance, can inadvertently
hinder reproducibility by imposing strict page or word limits that leave no room
for the necessary technical documentation, discouraging thoroughness in the name
of brevity.

Even when such instructions are given, they often fail across different
platforms or versions of the same platform. This lack of portability not only
hinders reproducibility but also complicates everyday research workflows.
Researchers using multiple machines must recreate environments consistently, and
collaborators must share identical computational setups to avoid
inconsistencies.

Finally, once the execution environment is correctly configured, additional
clarity is needed on how to *run* the project itself. Which packages should be
loaded first? Which scripts should be executed, and in what order? Without clear
documentation or automated orchestration, these operational details become yet
another barrier to reproducibility.

This paper focuses on two critical but often overlooked aspects of
reproducibility: *computational environment management* and *workflow
orchestration*. We present a comprehensive framework addressing both challenges
using the [Nix]{.pkg} package manager [@dolstra2004nix], making it accessible to
researchers through two [R]{.proglang} packages: [rix]{.pkg} and
[rixpress]{.pkg}. Before introducing these packages, we survey existing tools and
their limitations to contextualise our contributions.

A range of tools now exist to help researchers approach the gold standard of
full reproducibility, or to consistently deploy the same development environment
across multiple machines. Let us first consider the most basic step in this
process: listing the software used. In [R]{.proglang}, the `sessionInfo()`
function provides a concise summary of the software environment, including the R
version, platform details, and all loaded packages. Its output can be saved to a
file and included as part of a study’s reproducibility record. Below is an
example output from `sessionInfo()`:

::: {.content-hidden when-format="pdf"}
```R
sessionInfo()
```
:::

::: {.content-hidden when-format="pdf"}
```R
R version 4.3.2 (2023-10-31)
Platform: aarch64-unknown-linux-gnu (64-bit)
Running under: Ubuntu 22.04.3 LTS

Matrix products: default
BLAS:   /usr/lib/aarch64-linux-gnu/openblas-pthread/libblas.so.3
LAPACK: /usr/lib/aarch64-linux-gnu/openblas-pthread/libopenblasp[...]

locale:
 LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
 LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
 ...

attached base packages:
 stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
 nnet_7.3-19  mgcv_1.9-0   nlme_3.1-163

loaded via a namespace (and not attached):
 compiler_4.3.2  Matrix_1.6-1.1  tools_4.3.2     splines_4.3.2
 grid_4.3.2      lattice_0.21-9
```
:::

```{=tex}
\begin{CodeInput}
R> sessionInfo()
\end{CodeInput}
\begin{CodeOutput}
R version 4.3.2 (2023-10-31)
Platform: aarch64-unknown-linux-gnu (64-bit)
Running under: Ubuntu 22.04.3 LTS

Matrix products: default
BLAS:   /usr/lib/aarch64-linux-gnu/openblas-pthread/libblas.so.3
LAPACK: /usr/lib/aarch64-linux-gnu/openblas-pthread/libopenblasp[...]

locale:
 LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
 LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
 ...

attached base packages:
 stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
 nnet_7.3-19  mgcv_1.9-0   nlme_3.1-163

loaded via a namespace (and not attached):
 compiler_4.3.2  Matrix_1.6-1.1  tools_4.3.2     splines_4.3.2
 grid_4.3.2      lattice_0.21-9
\end{CodeOutput}
```

When an author includes this information, others attempting to reproduce the
study (future readers, collaborators, or even the author at a later time) can
see which version of [R]{.proglang} and which packages (with their versions)
were used. However, reproducing the environment still requires manually
installing the correct package versions --- a process that becomes difficult when
packages depend on system-level libraries. For example, the [sf]{.pkg} package
requires [GDAL]{.pkg}, [GEOS]{.pkg}, and [PROJ]{.pkg} libraries. Installing and
configuring these dependencies varies substantially across operating systems and
can be prohibitively difficult on some platforms.

A more robust approach than simply listing package versions is to use tools that
define per-project environments. This is common for [Python]{.proglang} users,
who have many tools at their disposal, such as the built-in *virtual
environments* module, or third-party tools like [uv]{.pkg}, which manages both
project-specific interpreters and package libraries using a lock file mechanism.

In the [R]{.proglang} ecosystem, virtual environments are not built-in as they
are in [Python]{.proglang}, however tools like [renv]{.pkg} provide equivalent
functionality. [renv]{.pkg} captures the project’s software state and writes it
to a lock file (`renv.lock`), which includes the exact versions of
[R]{.proglang} and all required packages. This lockfile serves as a blueprint
for restoring the environment automatically, ensuring others can recreate the
same setup with minimal effort.

However, both [renv]{.pkg} and [uv]{.pkg} have limitations. While [uv]{.pkg}
manages [Python]{.proglang} interpreter versions directly, [renv]{.pkg} does not
restore the version of [R]{.proglang} itself—installing the correct version must
be done separately using tools like [rig]{.pkg} [@rlib2023]. More critically,
neither handles system-level dependencies. If [sf]{.pkg} requires [GDAL]{.pkg}
version 3.0 but the system has version 2.4 installed, neither tool can resolve
this conflict. Users must manually install system libraries, and the process
differs across operating systems.

Pre-compiled binaries simplify installation but do not eliminate the need for
system libraries required at runtime. Other [R]{.proglang} packages such as
[groundhog]{.pkg} [@simonsohn2023] and [rang]{.pkg} [@chan2023] enable
date-based package installation, while the Posit Package Manager provides dated
CRAN snapshots. However, like [renv]{.pkg}, these focus on R packages and do not
address system-level dependencies or R version management.

These tools represent significant progress in reproducibility within their
ecosystems, but they share a common limitation: none handles system-level
dependencies or polyglot environments. A project using [R]{.proglang},
[Python]{.proglang} [@python3], and system tools like [Quarto]{.pkg} or \LaTeX
requires coordinating multiple package managers ([renv]{.pkg} for R, [uv]{.pkg}
for [Python]{.proglang}, manual installation for system tools), each with its own
configuration and potential conflicts.

The most comprehensive approach to date is containerisation with [Docker]{.pkg}.
[Docker]{.pkg} packages a *data product* together with all its
dependencies—including the operating system, system libraries, programming
languages, and packages—into a self-contained image. Once built, the analysis
can be executed inside a *container*, a running instance of the image.

The strength of [Docker]{.pkg} lies in its ability to bundle not only the
correct versions of [R]{.proglang} and its packages but also system-level
dependencies. Since [Docker]{.pkg} images are minimal Linux systems, they
include the libraries required by packages such as [sf]{.pkg}, ensuring that
future replicators have access to the same environment. Moreover, these images
can be shared via registries like Docker Hub, enabling reproducibility across
systems.

The Rocker project [@boettiger2017] provides pre-built [Docker]{.pkg} images for
the [R]{.proglang} community, offering specific R versions and, in some cases,
preinstalled packages. These serve as convenient bases for reproducible analysis
environments.

Despite its strengths, [Docker]{.pkg} presents several challenges:

1. Poor interactive support: While containers can be modified at runtime,
   changes are lost unless explicitly saved. Running graphical applications like
   RStudio Desktop requires complex X11 configuration that works reliably only
   on Linux. Web-based IDEs such as RStudio Server are alternatives but have
   their own limitations.

2. Steep learning curve: Writing reliable `Dockerfile`s requires familiarity
   with Linux system administration. To ensure reproducibility, `Dockerfile`s
   should reference image *digests* rather than *tags*, but this is rarely done.
   Packages like [dockerfiler]{.pkg} [@dockerfiler] help automate this but do
   not remove the need for basic Linux knowledge.

3. Post-hoc reproducibility: Analyses are often developed interactively and
   containerised only afterward, achieving post-hoc reproducibility rather than
   consistent environments during development.

4. Single-language focus: While [Docker]{.pkg} can support polyglot setups,
   orchestrating them requires manual coordination. There is no standard way to
   specify that one step runs in [Python]{.proglang} and another in [R]{.proglang}.

5. Incomplete reproducibility: Reproducibility is best viewed as a spectrum
   rather than a binary property [@dellaieraMasterThesis2024]. [Docker]{.pkg}
   achieves strong run-time reproducibility—consistency across systems
   (space)—but weak build-time reproducibility, or consistency over time.
   Mutable base images and non-deterministic commands such as `apt-get update`
   introduce temporal drift: rebuilding the same `Dockerfile` at different times
   can yield different results. Even with pinned base images, container
   checksums can vary across builds and machines. Thus, while [Docker]{.pkg}
   provides strong run-time reproducibility, achieving deterministic artefacts
   requires pinning image digests, freezing dependencies, and using
   deterministic build systems such as [Nix]{.pkg} for full control over both
   time and space [@Malka_2024].

Despite these shortcomings, [Docker]{.pkg} remains valuable for specific use
cases: production deployment, CI/CD integration, and cross-platform
distribution. The two approaches are not mutually exclusive. [Nix]{.pkg} can be
installed and used inside [Docker]{.pkg}, combining [Nix]{.pkg}'s deterministic
builds with [Docker]{.pkg}'s deployment infrastructure. In this workflow,
[Nix]{.pkg} handles environment reproducibility during the build of the image,
while [Docker]{.pkg} serves as the distribution mechanism. This addresses
[Docker]{.pkg}'s build-time reproducibility issues while maintaining its deployment
advantages. However, for research workflows where interactive development is
primary and [Docker]{.pkg}'s complexity is a barrier, [Nix]{.pkg} alone may suffice.

Thankfully, it is entirely possible to combine both [Docker]{.pkg} and [Nix]{.pkg}:
[Nix]{.pkg} to setup reproducible environments and [Docker]{.pkg} for deployment
across the many different already existing infrastructures.

Even with a perfectly reproducible environment, researchers face another
challenge: reliably executing the analysis workflow. Which scripts run first?
What are the dependencies between steps? Manual execution is error-prone and
poorly documented. Build automation tools like [Make]{.pkg} address this by
defining analyses as ordered, reproducible steps.

Within [R]{.proglang}, the [targets]{.pkg} package [@landau2021] provides a
modern, declarative approach to workflow management. It tracks dependencies,
caches results, and only recomputes affected steps, improving efficiency for
complex analyses. However, [targets]{.pkg} operates within a single
[R]{.proglang} session. For truly polyglot pipelines—where steps may require
incompatible dependencies or different language versions—this becomes
problematic. Running [targets]{.pkg} inside a [Docker]{.pkg} container ensures
reproducibility but forces the entire pipeline to share one environment.

@mcdermott2021 provides an excellent example of best practice. The accompanying
repository^[See
[https://github.com/grantmcdermott/skeptic-priors](https://github.com/grantmcdermott/skeptic-priors)]
demonstrates:

- Package versions recorded in an `renv.lock` file;
- A `Makefile` automating the full analysis;
- A `Dockerfile` providing the computational environment.

Achieving this required mastering multiple complex tools: [renv]{.pkg} for
package management, [Docker]{.pkg} for containerisation, and [Make]{.pkg} for
orchestration. The complexity increases for polyglot projects, which must manage
multiple ecosystems simultaneously.

Researchers therefore need a tool that:

1. Manages complete environments including languages, packages, and system
   dependencies;
2. Supports multiple languages natively;
3. Works interactively without container complexity;
4. Provides step-level isolation for different environments;
5. Ensures bit-for-bit reproducibility through deterministic builds;
6. Remains simple enough for researchers without systems administration
   expertise.

The [Nix]{.pkg} package manager provides these capabilities. It ensures
reproducible installation by deploying *component closures*—packages bundled
with all their direct and transitive dependencies [@dolstra2004nix]. Think of
installing a package as packing for a trip: traditional package managers assume
you’ll find essentials at your destination, while [Nix]{.pkg} packs everything
you need. This produces self-contained software environments that behave
identically across machines and over time.

[Nix]{.pkg} can replace [Docker]{.pkg} for isolation, [renv]{.pkg} for package
management, and even [Make]{.pkg} for orchestration—all within a single, unified
framework. However, it has a steep learning curve, with its own functional
language for declaratively defining software builds and configurations. While
this ensures reproducibility, it also creates a barrier to adoption.

To make [Nix]{.pkg} accessible to researchers, we developed two [R]{.proglang}
packages:

- [rix]{.pkg} generates [Nix]{.pkg} expressions from intuitive R function calls,
  eliminating the need to learn the [Nix]{.proglang} language. It handles
  environment definitions including [R]{.proglang}, [Python]{.proglang},
  [Julia]{.proglang} [@bezanson2017julia], system tools, and dependencies.

- [rixpress]{.pkg} orchestrates polyglot analytical pipelines using [Nix]{.pkg}
  as the build engine. Each step runs in its own hermetic environment with
  automatic caching and dependency tracking, enabling true polyglot workflows
  across [R]{.proglang}, [Python]{.proglang}, and [Julia]{.proglang}.
  [Python]{.proglang} users can use [ryxpress]{.pkg}, a direct port.

Together, these packages provide a single framework for managing reproducible,
polyglot environments and executing complex workflows with step-level isolation.
While [Nix]{.pkg} remains complex for advanced use, [rix]{.pkg} and
[rixpress]{.pkg} abstract this complexity, making deep reproducibility
accessible without systems administration expertise.

The remainder of this paper proceeds as follows. Section 2 introduces
[Nix]{.pkg} and explains how its functional model enables reproducibility.
Section 3 presents [rix]{.pkg} and demonstrates environment definition. Section
4 discusses the `rstats-on-nix` fork of the package repository. Section 5
introduces [rixpress]{.pkg} and demonstrates polyglot pipeline orchestration.
Section 6 concludes with discussion of future directions.

## The Nix package manager {#sec-nix}

[Nix]{.pkg} is a powerful, cross-platform package manager designed for
installing and building software in a fully reproducible and reliable manner.
Unlike traditional package managers, [Nix]{.pkg} emphasises immutable
infrastructure and a functional approach, which significantly enhances the
consistency of computational environments. As of writing, the primary [Nix]{.pkg}
package collection, `nixpkgs`, provides access to over 120,000 packages,
including nearly all of CRAN and Bioconductor. According to Repology, a service
that tracks package repositories across distributions, `nixpkgs` is both the
largest and among the most up-to-date package repositories available (see Figure
@fig-repology).

![Package repository size and freshness according to Repology. `nixpkgs` leads
in both total package count and update recency, making it well-suited for
reproducible research environments.](images/repology.png){#fig-repology}

This extensive coverage allows users to install not only [R]{.proglang} itself
but also all its required packages and system-level dependencies for any given
project.

While [Nix]{.pkg} is the default package manager for the NixOS Linux
distribution, it can be installed as a standalone tool on other Linux
distributions, macOS, and effectively on Windows via the Windows Subsystem for
Linux (WSL2), treating these as equivalent platforms in practice.

A key advantage of using [Nix]{.pkg} to install [R]{.proglang} packages, as
opposed to the standard `install.packages()` function, lies in its comprehensive
dependency management. [Nix]{.pkg} ensures that all dependencies of each package
are installed, irrespective of whether they are other [R]{.proglang} packages or
underlying system libraries.

Consider, for example, the [sf]{.pkg} package, which is fundamental for spatial
data analysis in [R]{.proglang}. [sf]{.pkg} relies on several complex external
system libraries such as [GDAL]{.pkg}, [GEOS]{.pkg}, and
[PROJ]{.pkg}. Manually installing and configuring these libraries to be
compatible with [sf]{.pkg} can be challenging and platform-specific. With
[Nix]{.pkg}, however, the user merely needs to declare [sf]{.pkg} as a project
requirement. [Nix]{.pkg} then automatically installs and configures all
necessary system libraries as transitive dependencies. This seamless process is
possible because the maintainers of the [R]{.proglang} programming language for
[Nix]{.pkg} have declaratively specified these dependencies for [sf]{.pkg} in
`nixpkgs`, allowing the entire software stack to be provisioned effortlessly
from an end-user perspective.

This concept is central to [Nix]{.pkg}'s design, where packages are referred to
as *component closures*. As @dolstra2004nix explains:

> The idea is to always deploy component closures: if we deploy a component,
> then we must also deploy its dependencies, their dependencies, and so on. That
> is, we must always deploy a set of components that is closed under the
> ''depends on'' relation. Since closures are selfcontained, they are the units
> of complete software deployment. After all, if a set of components is not
> closed, it is not safe to deploy, since using them might cause other
> components to be referenced that are missing on the target system.

The core of [Nix]{.pkg}'s reproducibility lies in its functional package
management paradigm and its unique approach to builds. When installing software
with [Nix]{.pkg}, it evaluates an *expression* written in the [Nix]{.proglang}
language. These expressions define *derivations*, which are declarative
blueprints describing how to build a package. A derivation specifies all inputs:
source code, build commands, and crucially, all direct and transitive
dependencies. This declarative nature ensures that every build process is fully
specified and independent of the environment where it's executed. Furthermore,
all [Nix]{.pkg} expressions for official packages are hosted in the `nixpkgs`
GitHub repository. By referencing a specific commit of `nixpkgs` (a process
known as *pinning a revision*) users guarantee that all packages installed are
derived from the exact same set of build instructions. This means that if an
environment is built today using a pinned revision, it will produce the
identical set of software versions and configurations if rebuilt tomorrow, next
year, or on a different machine, as long as the `nixpkgs` repository remains
accessible.

Pinning is essential, but it is not the only mechanism through which [Nix]{.pkg}
achieves reproducibility. [Nix]{.pkg} adopts principles from functional
programming, conceptualising software builds as *pure functions*. In
mathematics, a pure function is defined as one that always produces the same
output for a given input—for instance, if \( f(x) = x^2 \), then \( f(2) = 4 \)
invariably holds. Analogously, [Nix]{.pkg} ensures that, given identical inputs
(such as source code, dependencies, and build instructions), the resulting
output (the built package) remains identical, irrespective of the environment or
time of execution. In this sense, [Nix]{.pkg} functions as a *functional package
manager*, applying the deterministic properties of pure functions to software
construction.

This is achieved by disallowing side effects and ensuring that builds operate
within hermetic (isolated) environments, with no hidden dependencies on global
system state. The principles of functional package management (hermetic builds,
immutable artifacts, and declarative dependencies) are designed to produce
deterministic outputs. The effectiveness of this model was recently validated in
the most extensive study of build reproducibility to date. The effectiveness of
this model was recently validated at unprecedented scale by @malka2025, who
rebuilt 709,816 packages from historical nixpkgs snapshots (2017–2023). They
found rebuildability rates exceeding 99% and bitwise reproducibility rates
between 69% and 91%, with an upward trend over time. Importantly, they
identified that about 15% of unreproducible builds failed due to embedded build
timestamps, which a solvable issue rather than a fundamental limitation.

While this functional approach greatly enhances reproducibility, it can
sometimes introduce complexity for package maintainers, especially for software
that needs to download external assets during installation (e.g., certain
Bioconductor packages). For end-users, however, this complexity is largely
abstracted away thanks to both [rix]{.pkg} and [rixpress]{.pkg}.

Additionally, [Nix]{.pkg} supports multiple versions or *variants* of a package
on the same system. [Nix]{.pkg} installs all packages into a centralised,
content-addressed directory called the [Nix]{.pkg} store (typically
`/nix/store`). Each package gets a unique subdirectory whose name includes a
hash of all its inputs: source code, dependencies, build instructions. This
means `/nix/store/abc123-dplyr-1.0.0` and `/nix/store/xyz789-dplyr-1.1.0` can
coexist without conflicts.

This isolation ensures that projects always use their intended software
versions, allowing multiple versions of [R]{.proglang} to coexist and preventing
"dependency hell" and global environment pollution.

For a more in-depth technical discussion of [Nix]{.pkg}'s design principles, see
@dolstra2004nix.

By leveraging these principles, [Nix]{.pkg} can effectively replace and
integrate the functionalities of tools like [renv]{.pkg} and [Docker]{.pkg} for
[R]{.proglang} projects, or `requirements.txt` files and virtual environments
for [Python]{.proglang} projects. Furthermore, [Nix]{.pkg} excels at building
polyglot environments, which can seamlessly combine [R]{.proglang},
[Python]{.proglang}, [Julia]{.proglang}, a [LaTeX]{.proglang} distribution, or
any of the numerous other tools available in `nixpkgs`. This enables the
creation of a truly complete, project-specific, and deeply reproducible
environment that can be used interactively for development or non-interactively
for automated analysis. As long as the `nixpkgs` repository (or a suitable fork,
as discussed in Section @sec-fork) remains accessible, the environment can be
rebuilt reliably in the future.

[Nix]{.pkg} provides strong cross-platform support across major operating
systems. It runs natively on Linux (x86_64) with full functionality and optimal
integration. On macOS (Intel and Apple Silicon), however, reproducibility is
less reliable due to dependencies on impure system frameworks and Xcode
toolchains outside the Nix store. As a result, macOS users may need to update
project pins after system or Xcode upgrades to restore builds. On Windows,
[Nix]{.pkg} works well through the Windows Subsystem for Linux (WSL2), though
graphical applications may need extra setup. [Nix]{.pkg} environments on WSL can
be used interactively via VS Code or Positron for a smooth development
experience. It is also important to note that [Nix]{.pkg} can be installed
within a [Docker]{.pkg} image to deploy and reproduce environments reliably in
containerized workflows.

While [Nix]{.pkg} is a powerful system, it comes with notable challenges. The
steepest is its learning curve: [Nix]{.pkg} uses its own declarative language
(also called [Nix]{.proglang}), which can be difficult for those new to
functional or declarative paradigms. Complexity becomes most apparent when
writing custom derivations or troubleshooting intricate build issues. Build
times can also be long, especially for large environments, when pre-built
binaries aren’t available. Additionally, because [Nix]{.pkg} stores all package
outputs immutably in the Nix store, disk usage tends to be higher than with
traditional package managers. Finally, while `nixpkgs` is extensive, some new or
niche packages may be missing or fail to build across platforms.

To make these capabilities more approachable for [R]{.proglang} users seeking
reproducible, project-specific environments without learning the full
[Nix]{.proglang} language, we created the [rix]{.pkg} and [rixpress]{.pkg}
packages.

## Reproducible development environments with Nix {#sec-repro-nix}

As mentioned, [Nix]{.pkg} expressions are written in the [Nix]{.proglang}
programming language, which is purely functional. Here is a simple example that
creates a shell environment containing version 4.3.1 of [R]{.proglang}:

::: {.content-hidden when-format="pdf"}
```nix
let
  pkgs = import (fetchTarball
    "https://github.com/NixOS/nixpkgs/archive/976fa336.tar.gz"
  ) {};
  system_packages = builtins.attrValues {
    inherit (pkgs) R;
  };
in
  pkgs.mkShell {
    buildInputs = [ system_packages ];
    shellHook = "R --vanilla";
  }
```
:::

```{=tex}
\begin{CodeInput}
let
  pkgs = import (fetchTarball
    "https://github.com/NixOS/nixpkgs/archive/976fa336.tar.gz"
  ) {};
  system_packages = builtins.attrValues {
    inherit (pkgs) R;
  };
in
  pkgs.mkShell {
    buildInputs = [ system_packages ];
    shellHook = "R --vanilla";
  }
\end{CodeInput}
```

In this expression, the `let` keyword is used to define variables. The variable
`pkgs` imports the set of packages from the `nixpkgs` repository at the
specified commit `976fa336`. The variable `system_packages` lists the packages
to include in the environment; in this case, it is just the [R]{.proglang}
programming language, along with all its dependencies and their transitive
dependencies. The `mkShell` function then creates a development shell with the
specified packages. The `shellHook` is set to `"R --vanilla"`, meaning that
entering the shell automatically starts [R]{.proglang} in vanilla mode, ignoring
any startup options.

This expression can be saved in a file called `default.nix`. The environment can
then be built on a system with [Nix]{.pkg} installed using the `nix-build`
command.^[For installing [Nix]{.pkg}, we recommend the Determinate Systems
installer: \url{https://determinate.systems/posts/determinate-nix-installer}].
Once the build completes, the user can enter the interactive shell with
`nix-shell`, which executes following steps:

- Nix reads the expression and identifies what needs to be installed;
- It checks if these packages already exist in `/nix/store`;
- If not, it downloads or builds them (with all dependencies);
- It creates a new shell environment and updates the `$PATH` to include the
  right binaries from `/nix/store`.
- The user is now in an interactive [Nix]{.pkg} shell session.

This shell contains all the packages specified in `default.nix` and
can be used for development, similar to activating a virtual environment in the
[Python]{.proglang} ecosystem.

Writing [Nix]{.pkg} expressions can be challenging for users unfamiliar with the
[Nix]{.proglang} language. However, the ability to define a fully reproducible
development environment in a single text file and then rebuild it anywhere is
highly appealing. [rix]{.pkg} aims to lower the barrier to adoption of [Nix]{.pkg} for
reproducibility.

[rix]{.pkg} provides the `rix()` function, which simplifies generating
[Nix]{.pkg} expressions. It is available on CRAN and can be installed like any
other R package. Additionally, it can bootstrap an [R]{.proglang} development
environment on a system where [R]{.proglang} is not yet installed but
[Nix]{.pkg} is available. This can be done by running (inside of a terminal):

::: {.content-hidden when-format="pdf"}
```bash
nix-shell -I \
  nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
  R rPackages.rix
```
:::

```{=tex}
\begin{CodeInput}
$> nix-shell -I \
+    nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
+    R rPackages.rix
\end{CodeInput}
```

(the `-I` flag allows one to pass a specific revision of `nixpkgs`, ensuring
temporary shells are also reproducible).

This command opens a temporary [R]{.proglang} session with [rix]{.pkg}
available.^[`nix-shell -p` starts an interactive shell with the specified
packages.] From there, users can generate new [Nix]{.pkg} expressions for
building environments. For example, the following generates a `default.nix` file
that installs [R]{.proglang} 4.3.1 along with the [dplyr]{.pkg} and
[chronicler]{.pkg} packages:

::: {.content-hidden when-format="pdf"}
```r
library('rix')

rix(r_ver = "4.3.1",
  r_pkgs = c("dplyr", "chronicler"),
  project_path = ".",
  overwrite = TRUE)
```
:::

```{=tex}
\begin{CodeInput}
R> library('rix')

R> rix(r_ver = "4.3.1",
+    r_pkgs = c("dplyr", "chronicler"),
+    project_path = ".",
+    overwrite = TRUE)
\end{CodeInput}
```

[rix]{.pkg} can also handle more complex setups, and users can provide a date
instead of a specific [R]{.proglang} version:

::: {.content-hidden when-format="pdf"}
```r
rix(date = "2025-10-20",
  r_pkgs = c("rix", "dplyr", "chronicler", "AER@1.2-8"),
  system_pkgs = c("quarto", "git"),
  tex_pkgs = c(
        "amsmath",
        "framed",
        "fvextra",
        "environ",
        "fontawesome5",
        "orcidlink",
        "pdfcol",
        "tcolorbox",
        "tikzfill"
  ),
  git_pkgs = list(
      package_name = "fusen",
      repo_url = "https://github.com/ThinkR-open/fusen",
      commit = "60346860111be79fc2beb33c53e195f97504a667"
  ),
  ide = "positron",
  project_path = ".",
  overwrite = TRUE)
```
:::

```{=tex}
\begin{CodeInput}
R> rix(date = "2025-10-20",
+    r_pkgs = c("rix", "dplyr", "chronicler", "AER@1.2-8"),
+    system_pkgs = c("quarto", "git"),
+    tex_pkgs = c(
+          "amsmath",
+          "framed",
+          "fvextra",
+          "environ",
+          "fontawesome5",
+          "orcidlink",
+          "pdfcol",
+          "tcolorbox",
+          "tikzfill"
+    ),
+    git_pkgs = list(
+        package_name = "fusen",
+        repo_url = "https://github.com/ThinkR-open/fusen",
+        commit = "60346860111be79fc2beb33c53e195f97504a667"
+    ),
+    ide = "positron",
+    project_path = ".",
+    overwrite = TRUE)
\end{CodeInput}
```

This call to `rix()` generates a `default.nix` file for a development shell that
encapsulates a complete and reproducible research environment. It provides
[R]{.proglang} and its packages (including [AER]{.pkg} at version 1.2-8),
several TeXLive packages for \LaTeX document authoring, development versions of
[rix]{.pkg} and [fusen]{.pkg} pulled directly from GitHub at a specific commit,
and the Positron editor. The reproducibility of this environment is guaranteed
by pinning all components to a single point in time: the [R]{.proglang} packages
are resolved from the CRAN snapshot of October 20, 2025, while all other system
tools are fixed to the version of `nixpkgs` from that same date. Typing
`nix-shell` in a terminal within the folder that contains this `default.nix`
will drop the user into a new shell. It should be noted that it uses the
`nixpkgs` from that same date. It is also possible to configure IDEs to
dynamically load this new shell to provide the proper development tooling.

It is also possible to include [Python]{.proglang} and [Julia]{.proglang}
packages using the `py_conf` and `jl_conf` arguments respectively.
[Julia]{.proglang} packages work the same way as [R]{.proglang} packages: they
are pinned to their versions as of the specified date.

[Python]{.proglang} packages, however, require special handling. Unlike CRAN or
the Julia package registry, PyPI (the [Python]{.proglang} Package Index)
contains over 500,000 packages with highly variable quality—many lack proper
build specifications, have broken dependencies, or are abandoned. Because of
this heterogeneity, [Nix]{.pkg} maintainers cannot automatically package all of
PyPI. Instead, they manually curate and package individual [Python]{.proglang}
packages that meet quality standards. This means some [Python]{.proglang}
packages, or specific versions, may not be available through [Nix]{.pkg}.

For projects requiring [Python]{.proglang} packages not yet available in
[Nix]{.pkg}, [rix]{.pkg} supports integration with [uv]{.pkg}, a modern
[Python]{.proglang} package manager. This allows users to combine Nix's
system-level reproducibility ([Python]{.proglang} interpreter, system libraries)
with [uv]{.pkg}'s comprehensive package availability. While this hybrid approach
sacrifices some of Nix's deterministic build guarantees for [Python]{.proglang}
packages, it provides a pragmatic solution when the needed packages are
unavailable in `nixpkgs`.

Users can include [uv]{.pkg} as a system package:

::: {.content-hidden when-format="pdf"}
```r
rix(date = "2025-10-20",
  r_pkgs = c("rix", "dplyr", "chronicler"),
  system_pkgs = c("uv"),
  project_path = ".",
  overwrite = TRUE)
```
:::
```{=tex}
\begin{CodeInput}
R> rix(date = "2025-10-20",
+    r_pkgs = c("rix", "dplyr", "chronicler"),
+    system_pkgs = c("uv"),
+    project_path = ".",
+    overwrite = TRUE)
\end{CodeInput}
```

This hybrid approach offers complementary strengths: [Nix]{.pkg} manages the
[Python]{.proglang} interpreter, system libraries (e.g., [GDAL]{.pkg},
[HDF5]{.pkg}), and [R]{.proglang} packages with full determinism, while
[uv]{.pkg} handles [Python]{.proglang} packages through its own lockfile
mechanism (`uv.lock`). Within the Nix environment, users run standard [uv]{.pkg}
commands (e.g., `uv pip install pandas==2.0.0`) to install [Python]{.proglang}
packages. The [uv]{.pkg} lockfile records exact package versions and their
hashes, providing reproducibility for the [Python]{.proglang} package layer.

While this approach doesn't achieve Nix's level of build-time determinism for
[Python]{.proglang} packages ([uv]{.pkg} still downloads wheels or builds
packages at install time rather than using pre-built, content-addressed
artefacts) it provides a pragmatic balance: comprehensive [Python]{.proglang}
package availability with reasonable reproducibility guarantees. For polyglot
projects, this means [R]{.proglang} and system-level dependencies remain fully
reproducible through [Nix]{.pkg}, while [Python]{.proglang} packages gain the
reproducibility that [uv]{.pkg} can provide.

[rix]{.pkg} can generate [Nix]{.pkg} expressions even if [Nix]{.pkg} is not
installed on the system. This is useful for continuous integration and
continuous deployment (CI/CD) workflows on platforms such as GitHub Actions. For
instance, the repository containing the source code for this
article^[https://github.com/b-rodrigues/rix_paper] uses GitHub Actions to
compile the paper. Each time a push is made to the master branch, a runner
installs [Nix]{.pkg}, generates the environment from the hosted `default.nix`
file, and compiles the paper using [Quarto]{.pkg} within the reproducible
environment. This ensure that *exactly* the same environment is used on the
author's computer and on the CI/CD without any additional, platform-specific,
configuration.

Instead of first entering a [Nix]{.pkg} shell, it is also possible to run a
program directly from the environment:

::: {.content-hidden when-format="pdf"}
```bash
cd /path/to/project/ && nix-shell default.nix --run "Rscript analysis.R"
```
:::

```{=tex}
\begin{CodeInput}
cd /path/to/project/ && nix-shell default.nix --run "Rscript analysis.R"
\end{CodeInput}
```

This command runs `Rscript` and executes the `analysis.R` script, which in this
example should be located in the same directory as `default.nix`.

## The rstats-on-nix fork of nixpkgs {#sec-fork}

As explained earlier, [Nix]{.pkg} uses expressions from the `nixpkgs` GitHub
repository to build software. However, when generating expressions with
[rix]{.pkg}, the fork `rstats-on-nix/nixpkgs` is used instead.

Using a fork offers several advantages. First, it provides flexibility that the
official `nixpkgs` repository cannot always accommodate. [Nix]{.pkg} is
primarily the package manager for the NixOS Linux distribution, and governance
and technical choices made upstream can limit what [rix]{.pkg} aims to provide.

For instance, while [Nix]{.pkg} can theoretically support multiple versions (or
*variants*) of the same package, in practice maintainers cannot provide several
variants for all [R]{.proglang} packages, given the size of the ecosystem (over
20,000 CRAN and Bioconductor packages). This makes it difficult to install a
specific version of an [R]{.proglang} package not included in a particular
`nixpkgs` commit. With [rix]{.pkg}, users can install a specific package version
from source, e.g.:

::: {.content-hidden when-format="pdf"}
```r
rix(..., r_pkgs = "dplyr@1.0.7", ...)
```
:::

```{=tex}
\begin{CodeInput}
R> rix(..., r_pkgs = "dplyr@1.0.7", ...)
\end{CodeInput}
```

However, installing from source might fail, especially if the package needs to
be compiled.

Additionally, updating the full [R]{.proglang} package set on [Nix]{.pkg} daily
is impractical. While CRAN and Bioconductor update daily, the [R]{.proglang}
packages in `nixpkgs` are only updated with new [R]{.proglang} releases. This
limitation is due to [Nix]{.pkg}’s governance as a Linux distribution package
manager.

The `rstats-on-nix` fork allows me to circumvent these limitations. For example,
it provides daily snapshots of CRAN. Each day, the [R]{.proglang} package set is
updated and committed to a dated branch using GitHub Actions. Users can select a
specific date with:

::: {.content-hidden when-format="pdf"}
```r
rix(date = "2024-12-14", ...)
```
:::

```{=tex}
\begin{CodeInput}
R> rix(date = "2024-12-14", ...)
\end{CodeInput}
```

We strive to provide an available date per week: each Monday, a GitHub Action
tests popular packages on Linux and macOS, and only if all tests succeed is the
date added to the list of available dates in [rix]{.pkg}. This ensures users can
reliably install packages, and allows me to backport fixes if needed. For
example, when RStudio was temporarily broken due to a dependency issue
(`boost`), a pull request was submitted to the official `nixpkgs` repository. I
backported the fix to the `rstats-on-nix` fork, making RStudio available to
users of [rix]{.pkg} earlier than upstream, as merging PRs in the official
repository can take some time.

We have backported fixes to the `rstats-on-nix` `nixpkgs` fork as far back as
March 2019. The process involves checking out a `nixpkgs` commit on the selected
date, updating the [R]{.proglang} package set using Posit CRAN and Bioconductor
snapshots, backporting fixes, and ensuring popular packages work on both
x86-linux (including WSL2) and aarch64-darwin (Apple Silicon). These changes are
committed to a dated branch in `rstats-on-nix/nixpkgs`. Users can see all
available dates with `rix::available_dates()`.

A drawback of forking `nixpkgs` is that backported packages are not included
upstream and thus are not prebuilt by Nix’s CI platform, Hydra. Users may need
to build many packages from source, which can be time-consuming. To mitigate
this, we provide a binary cache sponsored by [Cachix](https://www.cachix.org/),
complementing the public Nix cache. Instructions for using Cachix are in
[rix]{.pkg}’s documentation. Using the cache significantly speeds up
installations, as prebuilt packages are downloaded rather than compiled.

The fork also allows me to catch issues (such as packages’ builds breaking) early
on, and prepare fixes that can then be contributed upstream.

While the reliance on the `rstats-on-nix` fork might initially raise
sustainability concerns, this risk should be contextualized. The fork serves a
focused purpose: it updates CRAN and [Julia]{.proglang} package registries daily
and backports fixes when builds break. Importantly, users are never locked into
this fork—they retain the ability to point directly to any commit in the
upstream `nixpkgs` repository, maintaining full access to the official
[Nix]{.pkg} ecosystem. The fork introduces no fundamental architectural changes
or proprietary extensions; it functions as a convenience layer providing tested
package sets and dated snapshots. Should maintenance cease, users could
seamlessly transition to upstream `nixpkgs` commits, though with trade-offs:
less frequent [R]{.proglang} and [Julia]{.proglang} package updates. The fork's
value lies not in creating dependency, but in reducing friction through testing
infrastructure and curated "known-good" dates. Thus, the fork represents an
enhancement to user experience rather than a critical single point of
failure—the underlying [Nix]{.pkg} infrastructure and official `nixpkgs`
repository remain the foundation, ensuring long-term viability of environments
created with [rix]{.pkg}.

## Orchestrating the workflow with rixpress {#sec-rixpress}

Defining a reproducible environment with [rix]{.pkg} addresses the first major
challenge of reproducibility. The second challenge is reliably and efficiently
executing the analysis workflow within that environment, which is the role of
the [rixpress]{.pkg} package (or [ryxpress]{.pkg} for [Python]{.proglang}
users).

As mentioned in the introduction, a build automation tool like [targets]{.pkg}
is invaluable for managing complex analyses. It tracks dependencies between code
and data, caches results, and only recomputes steps that have changed. One can
run a [targets]{.pkg} pipeline inside a Nix environment to make it reproducible.
However, this approach has limitations: the entire pipeline must run in a single
environment, and orchestrating steps across different languages (e.g.,
[R]{.proglang} and [Python]{.proglang}) requires manual handling via packages
like [reticulate]{.pkg}.

[rixpress]{.pkg} overcomes these limitations by using [Nix]{.pkg} not just as a
package manager, but as the build automation engine itself. In a
[rixpress]{.pkg} pipeline, each step is defined as a [Nix]{.pkg} derivation,
providing two key benefits:

  1. True Polyglot Pipelines: Each step can have its own [Nix]{.pkg} environment. A
     [Python]{.proglang} step can run in a pure [Python]{.proglang} environment,
     an [R]{.proglang} step in an [R]{.proglang} environment, and a Quarto
     rendering step in yet another, all within the same pipeline.
  2. Deep Reproducibility: Each step is a hermetically sealed [Nix]{.pkg} derivation
     whose output is cached in the [Nix]{.pkg} store based on the hash of all its
     inputs. All artefacts are direct children of the computational environment
     (because the computational environment is actually a dependency of the
     artefacts), ensuring they are rebuilt if the environment changes, keeping
     environment and outputs always in sync.

To illustrate these capabilities, we present a polyglot pipeline that simulates
a Real Business Cycle (RBC) model in [Julia]{.proglang}, uses the resulting data
to train an [XGBoost]{.pkg} forecasting model in [Python]{.proglang}, visualises
the results in [R]{.proglang} with [ggplot2]{.pkg}, and finally compiles a
[Quarto]{.pkg} report. The code for this example can be found in the [following
repository](https://github.com/b-rodrigues/rixpress_demos/tree/master/case-study)^[https://github.com/b-rodrigues/rixpress_demos/tree/master/case-study],
with additional details provided in the Appendix of this manuscript.

For new projects, [rixpress]{.pkg} provides `rxp_init()`, which generates the
necessary boilerplate structure including a pipeline definition file and
template helper scripts. Users then define their pipeline in the script
`gen-pipeline.R` using functions inspired by [targets]{.pkg}, and the
computational environment the pipeline will be executed in the `gen-env.R`
script. If functions are needed for pipeline steps, these are to be specified, in
language-specific helper scripts: `functions.jl` for [Julia]{.proglang},
`functions.py` for [Python]{.proglang}, and `functions.R` for [R]{.proglang}.
This separation keeps the pipeline definition clean and readable while
encapsulating the scientific logic in appropriate language-native files.
Importantly, these helper scripts are standard language files—they contain no
[rixpress]{.pkg}-specific code and can be reused in other projects or executed
independently outside of [Nix]{.pkg}.

The following example demonstrates how [rixpress]{.pkg} orchestrates a granular,
multi-step data science workflow that crosses language boundaries. (Complete
setup instructions and a detailed walkthrough are provided in Appendix)

::: {.content-hidden when-format="pdf"}
```r
library('rixpress')

pipeline_steps <- list(
  # STEP 0: Define RBC Model Parameters in Julia
  rxp_jl(alpha, 0.3),
  rxp_jl(beta, 1 / 1.01),
  # ... (other parameters omitted for brevity) ...

  # STEP 1: Julia - Simulate the RBC model
  rxp_jl(
    name = simulated_rbc_data,
    expr = "simulate_rbc_model(alpha, beta, delta, rho, sigma, sigma_z)",
    user_functions = "functions/functions.jl",
    encoder = "arrow_write"
  ),

  # STEP 2.1: Python - Prepare features
  rxp_py(
    name = processed_data,
    expr = "prepare_features(simulated_rbc_data)",
    user_functions = "functions/functions.py",
    decoder = "pyarrow.feather.read_feather"
  ),

  # STEP 2.2: Python - Split data (X_train, y_train, etc.)
  rxp_py(name = X_train, expr = "get_X_train(processed_data)", ...),
  # ... (other data splits omitted for brevity) ...

  # STEP 2.3: Python - Train the XGBoost model
  rxp_py(
    name = trained_model,
    expr = "train_model(X_train, y_train)",
    user_functions = "functions/functions.py"
  ),

  # STEP 2.4: Python - Make predictions and format results
  # ... (prediction and formatting steps omitted for brevity) ...
  rxp_py(
    name = final_predictions_df,
    expr = "format_results(y_test, model_predictions)",
    user_functions = "functions/functions.py",
    encoder = "save_arrow"
  ),

  # STEP 3: R - Visualise the predictions
  rxp_r(
    name = output_plot,
    expr = plot_predictions(final_predictions_df),
    user_functions = "functions/functions.R",
    decoder = arrow::read_feather
  ),

  # STEP 4: Quarto - Compile the final report
  rxp_qmd(
    name = final_report,
    qmd_file = "readme.qmd"
  )
)

# Generate the 'pipeline.nix' file from the R list
rxp_populate(pipeline_steps, build = TRUE)
```
:::

```{=tex}
\begin{CodeInput}
R> library('rixpress')

R> pipeline_steps <- list(
+  # STEP 0: Define RBC Model Parameters in Julia
+  rxp_jl(alpha, 0.3),
+  rxp_jl(beta, 1 / 1.01),
+  # ... (other parameters omitted for brevity) ...
+
+  # STEP 1: Julia - Simulate the RBC model
+  rxp_jl(
+    name = simulated_rbc_data,
+    expr = "simulate_rbc_model(alpha, beta, delta, rho, sigma, sigma_z)",
+    user_functions = "functions/functions.jl",
+    encoder = "arrow_write"
+  ),
+
+  # STEP 2.1: Python - Prepare features
+  rxp_py(
+    name = processed_data,
+    expr = "prepare_features(simulated_rbc_data)",
+    user_functions = "functions/functions.py",
+    decoder = "pyarrow.feather.read_feather"
+  ),
+
+  # STEP 2.2: Python - Split data (X_train, y_train, etc.)
+  rxp_py(name = X_train, expr = "get_X_train(processed_data)", ...),
+  # ... (other data splits omitted for brevity) ...
+
+  # STEP 2.3: Python - Train the XGBoost model
+  rxp_py(
+    name = trained_model,
+    expr = "train_model(X_train, y_train)",
+    user_functions = "functions/functions.py"
+  ),
+
+  # STEP 2.4: Python - Make predictions and format results
+  # ... (prediction and formatting steps omitted for brevity) ...
+  rxp_py(
+    name = final_predictions_df,
+    expr = "format_results(y_test, model_predictions)",
+    user_functions = "functions/functions.py",
+    encoder = "save_arrow"
+  ),
+
+  # STEP 3: R - Visualise the predictions
+  rxp_r(
+    name = output_plot,
+    expr = plot_predictions(final_predictions_df),
+    user_functions = "functions/functions.R",
+    decoder = arrow::read_feather
+  ),
+
+  # STEP 4: Quarto - Compile the final report
+  rxp_qmd(
+    name = final_report,
+    qmd_file = "readme.qmd"
+  )
+)

# Generate the 'pipeline.nix' file from the R list
R> rxp_populate(pipeline_steps, build = TRUE)
\end{CodeInput}
```

Computation steps, referred to as *derivations*, are defined using the functions
`rxp_r()`, `rxp_jl()`, and `rxp_py()`, corresponding to [R]{.proglang},
[Julia]{.proglang}, and [Python]{.proglang} environments, respectively. Variants
of these functions suffixed with `_file()` designate *import derivations*, which
introduce external data into the pipeline. It is important to note that any data
imported in this way is stored within the [Nix]{.pkg} store, ensuring
reproducibility but also persisting the data in the build environment. The
function `rxp_qmd()` is used to compile a [Quarto]{.pkg} document (there is also
`rxp_rmd()` which serves the same purpose for [R Markdown]{.pkg} documents).

The `rxp_populate()` function translates this [R]{.proglang} list into a
`pipeline.nix` file, which declaratively defines the entire workflow. Data flows
from derivation to derivation by being serialised into the efficient and
language-agnostic Arrow format using the pair of `encoder/decoder` functions.
Other universal formats, such as `csv` or `json` could have been used.

As a sidenote: [Python]{.proglang} users who wish to use [ryxpress]{.pkg} define
pipelines using the same [R]{.proglang}-based DSL shown above. This design
choice keeps the pipeline definition language consistent across both
[R]{.proglang} and [Python]{.proglang} ecosystems. [ryxpress]{.pkg} calls
[R]{.proglang} and [rixpress]{.pkg} under the hood to generate and build the
`pipeline.nix` file. To use [ryxpress]{.pkg}, one should simply add this
package to the `py_conf` argument of `rix()`.

![Graphical representation of the polyglot pipeline. Purple nodes are
[Julia]{.proglang}, yellow nodes are [Python]{.pkg}, light blue nodes are
[R]{.proglang}, and dark blue nodes are Quarto
derivations.](images/dag.png){#fig-dag}

The package can then generate a visual representation of the pipeline's directed
acyclic graph by running `rxp_dag()`, as can be seen in @fig-dag.

To execute the pipeline, one can either set `build = TRUE` in `rxp_populate()`
or call `rxp_make()` separately. [Nix]{.pkg} executes each step in order,
building dependencies as needed. Outputs are cached, so subsequent runs only
recompute steps with changed inputs or code. This provides the efficiency of
[targets]{.pkg} with polyglot support and bit-for-bit reproducibility. Artefacts
can be inspected interactively in [R]{.proglang} using
`rxp_read("artefact_name")` or `rxp_load("artefact_name")`.

[rixpress]{.pkg} also includes several additional features not covered here for
brevity. As mentioned previously, it is also possible to configure popular IDEs
to work interactively and seamlessly with both [rix]{.pkg} and [rixpress]{.pkg},
enabling a smooth, reproducible workflow from within the development
environment. Detailed setup instructions are provided in the vignettes of both
packages.

## A Comparison of Pipeline Paradigms: Imperative vs. Declarative {#sec-comparison}

To illustrate the practical benefits of our framework, we implemented the RBC
pipeline example from Section @sec-rixpress using two distinct paradigms. The
first is a traditional, imperative stack combining Docker for environment
isolation, Make for workflow orchestration, and language-specific package
managers. The second is our proposed declarative framework using `rix` and
`rixpress`. While both approaches successfully execute the same polyglot
analysis to produce identical outputs, their comparison illuminates fundamental
differences in conceptual models, system complexity, and the nature of the
reproducibility guarantees they provide.

### The Imperative Composition of Specialized Tools

The traditional approach represents a composition of specialized tools, each
solving part of the reproducibility puzzle. This paradigm is entirely
imperative; the researcher must explicitly script the procedural steps for both
environment construction and workflow execution.

The codification of the environment is a procedural build process specified in a
`Dockerfile`. Successfully authoring such a file necessitates that the
researcher act as a system administrator, manually scripting a sequence of shell
commands to install each component. This includes managing `apt` repositories
for R, invoking external installers like `uv` and `rig` for Python and R
runtimes, and using `curl` and `tar` for Julia. Subsequently, separate commands
must be run to install packages within each language's ecosystem. The procedural
verbosity and intertwined concerns of this approach are reflected in the
resulting `Dockerfile`, which spans 108 lines to define the complete
environment.

Here is a sketch of this `Dockerfile`:

```dockerfile
# Add R repository and install specific version
RUN apt-get update && apt-get install -y software-properties-common
RUN add-apt-repository ppa:...
RUN curl -L https://rig.r-pkg.org/... | sh
RUN rig add 4.5.1

# Install Python with uv
RUN curl -LsSf https://astral.sh/uv/install.sh | sh
RUN uv python install 3.13

# Download and extract Julia
RUN curl -fsSL https://julialang-s3.julialang.org/... -o julia.tar.gz
RUN tar -xzf julia.tar.gz -C /opt/

# Install packages for each language separately
RUN echo 'options(repos = c(CRAN = ...))' > /root/.Rprofile
RUN Rscript -e 'install.packages(...)'

# Install Python packages using uv with specific versions for reproducibility.
RUN echo "pandas==2.3.3" > /tmp/requirements.txt && \
    echo "scikit-learn==1.7.2" >> /tmp/requirements.txt && \
    # ... more packages ...

RUN uv pip install --no-cache -r /tmp/requirements.txt && \
    rm /tmp/requirements.txt

# Install specific versions of Julia packages for reproducibility
RUN julia -e 'using Pkg; \
    Pkg.add(name="Arrow", version="2.8.0"); \
    # ... more packages ...'
```

This procedural philosophy extends from environment definition to workflow
orchestration. A `Makefile` defines the pipeline as a directed acyclic graph of
file dependencies. While robust, this model requires the researcher to manually
encode the dependency graph, tightly coupling the analytical logic to specific
file paths. Furthermore, since `make` is agnostic to the content of these files,
it provides no mechanism to enforce type or schema consistency between steps.

A direct consequence of this file-based orchestration model is the proliferation
of wrapper scripts. Each step in the `Makefile` invokes a script containing a
significant amount of procedural boilerplate unrelated to the scientific logic.
This includes parsing command-line arguments, handling data serialization and
deserialization, and managing file paths. The core analytical functions are thus
obscured by this necessary but distracting scaffolding.

The cumulative effect is a system of high structural complexity. The traditional
implementation requires nine separate files to fully specify the environment,
workflow, and analysis. This includes the `Dockerfile`, `Makefile`, three
wrapper scripts for orchestration, and three `functions` files containing the
core logic, in addition to the final report. The cognitive overhead is
substantial, as a user must comprehend the conventions and interactions of this
entire toolchain.

### A Unified Declarative Framework

The Nix-based framework inverts this imperative model. Instead of scripting the
procedural construction of the environment and workflow, the researcher declares
the desired end-state.

Environment specification is reduced to a single declarative statement in the
`gen-env.R` script. The researcher specifies *what* the environment should
contain, leaving the procedural implementation of *how* to build it entirely to
the Nix toolchain. The `gen-env.R` script replaces the 108-line `Dockerfile`,
abstracting away all system administration details. `rix` translates this
high-level specification into a formal Nix expression that deterministically
resolves all transitive dependencies for the complete polyglot environment.

```r
library(rix)
rix(
  date = "2025-10-14",
  r_pkgs = c("ggplot2", "dplyr", "arrow"),
  jl_conf = list(jl_version = "lts", ...),
  py_conf = list(py_version = "3.13", ...),
  ...
)
```

Workflow orchestration undergoes a similar conceptual shift from procedural to
declarative. The `gen-pipeline.R` script defines the pipeline not as a set of
file rules, but as a graph of data objects. Dependencies are not manually
specified but are automatically inferred by `rixpress` from the functional
relationships in the code. This object-centric model renders the procedural
wrapper scripts unnecessary. The scientific logic is expressed as a set of pure
functions that accept data structures as inputs and return them as outputs, free
of any boilerplate for file I/O or command-line parsing.

```python
# functions/functions.py
def train_and_predict_output(simulated_df: pd.DataFrame) -> pd.DataFrame:
    # ... scientific logic ...
    return format_results(predictions)
```

This declarative approach significantly reduces system complexity. The entire
implementation requires only six files: two for the high-level environment and
pipeline declarations, three containing the pure scientific functions, and the
final report. The reduction is not merely quantitative; the remaining files are
conceptually simpler, containing only analytical logic and specifications, with
all procedural scaffolding abstracted away by the framework.

### Reproducibility Guarantees: Space and Time

The critical distinction between the two paradigms lies in the nature of their
reproducibility guarantees. The traditional stack, anchored by Docker, provides
strong **spatial reproducibility**—the ability to execute the analysis
identically across different machines. However, it is vulnerable to **temporal
drift**, a concept that distinguishes reproducibility across space from
consistency over time [@malka2025space]. This temporal vulnerability arises from
several sources, including the resolution of mutable base image tags (e.g.,
`ubuntu:24.04`) and the potential for language-level package managers to resolve
dependencies differently over time.

The Nix-based approach is architected to solve this problem of temporal drift.
By pinning the entire software ecosystem to a single, immutable revision of the
`nixpkgs` repository, the framework provides strong guarantees of **temporal
reproducibility**. This claim is supported by large-scale empirical evidence; a
study by @malka2024yes, which rebuilt over 700,000 historical packages,
confirmed that Nix's functional model achieves a bit-for-bit reproducibility
rate exceeding 90% and a rebuildability rate over 99% across a multi-year
timeframe.

### Summary of Differences

| Dimension | Traditional Stack | Nix-Based Framework |
|-----------|------------------|---------------------|
| **Paradigm** | Imperative composition of specialized tools | Declarative specification in unified framework |
| **Environment setup** | Step-by-step shell commands | High-level specification |
| **Workflow definition** | File-based rules in Make syntax | Object-based graph in R syntax |
| **Scientific code** | Wrapped in CLI parsing and file I/O boilerplate | Pure functions with no plumbing |
| **File count** | 9 files (4 for orchestration and plumbing) | 6 files (2 for orchestration) |
| **Expertise required** | Linux system administration, Make syntax | R programming |
| **Spatial reproducibility** | Strong (via Docker isolation) | Strong (via Nix isolation) |
| **Temporal reproducibility** | Moderate (vulnerable to drift) | Strong (validated empirically [@malka2025]) |

This declarative rigor, however, is not without pragmatic costs. Our validation
on GitHub Actions shows that the imperative Docker-based runner completes the
entire pipeline in approximately three minutes, whereas the declarative
Nix-based runner takes around five minutes. This difference reflects the
trade-off between the upfront computational cost of Nix's rigorous dependency
resolution and the faster, but more brittle, imperative builds of the
traditional approach.

## Conclusion {#sec-conclusion}

Many tools exist to improve reproducibility, but [Nix]{.pkg} stands out because
it deploys complete software environments *closed under the "depends on"
relation*: it installs not only a package, but all its dependencies and their
dependencies. This makes [Nix]{.pkg} particularly powerful for reproducible
research.

However, solving such a complex problem makes [Nix]{.pkg} a complex tool. With
[rix]{.pkg}, we aim to make [Nix]{.pkg} more accessible to [R]{.proglang} users
by providing a familiar interface and workflow. By building reproducible
development shells with [Nix]{.pkg}, researchers can accommodate a wide range of
use cases: running scripts and pipelines, developing interactive [shiny]{.pkg}
applications, or serving [plumber]{.pkg} APIs.

Furthermore, [rixpress]{.pkg} extends this reproducibility to entire analysis
pipelines. By leveraging [Nix]{.pkg} as a build automation engine, [rixpress]{.pkg}
allows polyglot workflows where each step runs in its own hermetically sealed
environment. This ensures deep reproducibility, efficient caching, and seamless
orchestration of polyglot analyses. Most importantly, it provides native
polyglot support without the manual coordination required by containerization
approaches, making complex multi-language workflows accessible to researchers
without systems administration expertise.

Adopting [Nix]{.pkg}-based workflows represents a significant paradigm shift, and
researchers should approach it incrementally. A reasonable pathway begins with
using [rix]{.pkg} to generate reproducible environments for new projects, while
maintaining existing workflows for ongoing work. This allows learning [Nix]{.pkg}
concepts without risking active research. As comfort grows, researchers can
experiment with [rixpress]{.pkg} for smaller analyses before tackling complex
polyglot pipelines. Institutions can support adoption by providing workshops,
maintaining institutional binary caches with private packages, and designating
local experts. The investment pays increasing returns: early projects require
substantial learning, but subsequent projects leverage accumulated knowledge and
reusable environment definitions. For research groups, having one member develop
[Nix]{.pkg} expertise can enable the entire team to benefit from reproducible
environments without everyone climbing the learning curve. Ultimately, the
decision to adopt these tools depends on project timescales, reproducibility
requirements, and institutional context—but the framework exists for those who
need it.

While [rix]{.pkg} and [rixpress]{.pkg} (and [ryxpress]{.pkg}) provide
significant steps forward, it is important to acknowledge current limitations.
Debugging complex [rixpress]{.pkg} pipelines remains more difficult than
debugging [targets]{.pkg} workflows, particularly when failures occur deep in
the [Nix]{.pkg} build process. The visualization capabilities, while useful, lack
features like automatic detection of outdated derivations that make tools like
[targets]{.pkg} particularly developer-friendly. Cross-platform reproducibility,
while generally strong, has known limitations on macOS due to system framework
dependencies. These are not insurmountable problems—they represent areas for
continued development and community contribution. Researchers adopting these
tools should do so with realistic expectations: they provide powerful
capabilities for deep reproducibility and polyglot workflows, but require
patience during the learning phase and may occasionally present challenges that
simpler tools avoid.

## Acknowledgments {.unnumbered}

We thank the rOpenSci reviewers and contributors who provided valuable feedback
on the development of [rix]{.pkg} and [rixpress]{.pkg}. In particular, we are
grateful to David Watkins and Jacob Wujiciak-Jens for their reviews of
[rix]{.pkg}, and to William Landau and Anthony Martinez for their reviews of
[rixpress]{.pkg}. We also acknowledge the contributions of Richard J. Acton,
Jordi Rosell, Elio Campitelli, László Kupcsik, and Michael Heming for
[rix]{.pkg}. Their expertise and feedback greatly improved the quality and
usability of these packages. We would also like to thank Pol Dellaiera and Edvin
Syk for providing feedback on the manuscript.

## References {.unnumbered}

:::{#refs}

:::

## Appendix {#sec-appendix}

### Reproducing this paper

The source code of this paper is hosted on GitHub and can be found
at following
[link](https://github.com/b-rodrigues/rix_paper)^[https://github.com/b-rodrigues/rix_paper].
The paper can be easily compiled by running the following command:

::: {.content-hidden when-format="pdf"}
```bash
nix-shell --run "quarto render paper.qmd --to jss-pdf"
```
:::

```{=tex}
\begin{CodeInput}
$> nix-shell --run "quarto render paper.qmd --to jss-pdf"
\end{CodeInput}
```

The `default.nix` file defines the exact computational environment required to
compile the manuscript, ensuring full reproducibility of the build process. To
guarantee the reproducibility of the manuscript, it is automatically recompiled
via GitHub Actions upon each commit, using the same [Nix]{.pkg} environment. An
HTML version is additionally deployed to GitHub Pages, providing an accessible
format for viewing on smaller devices at this
[link](https://b-rodrigues.github.io/rix_paper/)^[[https://b-rodrigues.github.io/rix_paper/](https://b-rodrigues.github.io/rix_paper/)].
The PDF version of the manuscript can be found at
[link](https://b-rodrigues.github.io/rix_paper/paper.pdf)^[[https://b-rodrigues.github.io/rix_paper/paper.pdf](https://b-rodrigues.github.io/rix_paper/paper.pdf)].

### A Complete Polyglot Example with rixpress

This appendix provides a conceptual walk-through of the polyglot pipeline
example discussed in Section 5. The pipeline simulates a Real Business Cycle
(RBC) model in [Julia]{.proglang}, trains an [XGBoost]{.pkg} model in
[Python]{.proglang}, visualises the results in [R]{.proglang}, and compiles a
final report with [Quarto]{.pkg}.

The complete, runnable source code for this example is available in the paper's
GitHub repository. A shell script, `run_polyglot_example.sh`, is provided to
automate the entire process from environment creation to final output, as
described at the end of this
section^[\url{https://github.com/b-rodrigues/rix_paper/blob/master/repro-script/run_polyglot_example.sh}].

#### Project Structure and Components

The project is organized into a set of scripts, each with a distinct
responsibility. The core logic is separated from the environment definition and
pipeline orchestration.

*   `functions/`: This directory contains the helper scripts with the core analytical code for each language.
    *   `functions.jl`: The [Julia]{.proglang} script for the economic simulation.
    *   `functions.py`: The [Python]{.proglang} script for the machine learning workflow.
    *   `functions.R`: The [R]{.proglang} script for data visualisation.
*   `gen-env.R`: An [R]{.proglang} script that defines the reproducible computational environment.
*   `gen-pipeline.R`: The main [R]{.proglang} script that defines and orchestrates the entire polyglot pipeline.

#### Step 1: The Environment Definition

The foundation of the project is the reproducible environment, defined in
`gen-env.R`. This script uses the `rix()` function to programmatically generate
a `default.nix` file. This file serves as a complete blueprint for the
computational environment, specifying:

*   The exact versions of [R]{.proglang}, [Python]{.proglang}, and
    [Julia]{.proglang}.
*   A list of required packages for each language (e.g., [ggplot2]{.pkg} for
  [R]{.proglang}, [xgboost]{.pkg} for [Python]{.proglang}, [DataFrames]{.pkg}
  for [Julia]{.proglang}).
*   System-level dependencies like [Quarto]{.pkg}.

Crucially, the entire environment is pinned to a specific date (`2025-10-14`),
ensuring that anyone who builds the environment, now or in the future, will get
the exact same software versions, guaranteeing reproducibility.

By dropping into a temporary shell using the following command:

::: {.content-hidden when-format="pdf"}
```bash
nix-shell -I \
  nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
  R rPackages.rix
```
:::

```{=tex}
\begin{CodeInput}
$> nix-shell -I \
+  nixpkgs=https://github.com/rstats-on-nix/nixpkgs/tarball/2025-10-20 -p \
+  R rPackages.rix
\end{CodeInput}
```

it is possible to generate the `default.nix` by sourcing `gen-env.R`. Then, one
leaves this temporary shell, and builds the environment using the command
`nix-shell`, which also drop the user into the development shell.

#### Step 2: The Core Analytical Logic

The scientific logic for each stage of the pipeline is encapsulated in the
separate helper scripts within the `functions/` directory.

*   The RBC Model Simulation (`functions.jl`): This [Julia]{.proglang} script
  contains a single pure function that implements the state-space solution to
  the RBC model, based on @depaoli2009rbc. It takes the model's economic
  parameters as inputs and returns the simulated time-series data as a
  [DataFrame]{.pkg}.
*   The [XGBoost]{.pkg} Forecasting Model (`functions.py`): This
  [Python]{.proglang} script handles the complete machine learning workflow
  through a series of modular functions. Its responsibilities include feature
  engineering (creating lagged variables), splitting the data into training and
  testing sets, training the [XGBoost]{.pkg} model, and generating predictions.
*   The Visualisation (`functions.R`): This [R]{.proglang} script contains a
  function that uses [ggplot2]{.pkg} to create the final visualisation. It is
  designed to take the data frame of actual and predicted values produced by the
  [Python]{.proglang} script and generate a plot comparing the two series.

#### Step 3: Orchestrating the Pipeline

The entire workflow is defined and orchestrated by `gen-pipeline.R`. This script
acts as the master plan, using functions from the [rixpress]{.pkg} package to
define each computational step as a *derivation*. It declaratively outlines the
dependencies between steps:

1.  It begins by defining the RBC model parameters in [Julia]{.proglang}.
2.  It specifies that the RBC simulation in `functions.jl` depends on these
    parameters.
3.  It then defines the series of [Python]{.proglang} steps (feature
  preparation, training, prediction), making each one dependent on the output of
  the previous one. The first [Python]{.proglang} step is explicitly made
  dependent on the output from the [Julia]{.proglang} simulation.
  [rixpress]{.pkg} handles the passing of data between languages, in this case
  using the [Arrow]{.pkg} file format for efficiency.
4.  Finally, it defines the **R visualisation** step, which depends on the final
  predictions from the [Python]{.proglang} model, and a **Quarto rendering**
  step that depends on the generated plot.

When this script is run in the development shell (by executing
`source("gen-pipeline.R")`), [rixpress]{.pkg} translates the declared pipeline
into a master [Nix]{.proglang} expression that [Nix]{.pkg} can execute,
automatically handling the caching of results and re-running only the necessary
steps if a piece of code or data changes.

To build the pipeline from an interactive [Python]{.proglang} session, one would
execute the following lines:

::: {.content-hidden when-format="pdf"}
```python
from ryxpress import rxp_make
rxp_make()
```
:::

```{=tex}
\begin{CodeInput}
Python> from ryxpress import rxp_make
Python> rxp_make()
\end{CodeInput}
```

#### Running the Project

This entire example can be executed by running the `run_polyglot_example.sh`
script available in the root of the paper's repository.
This script automates the full process:
1.  It first executes `gen-env.R` inside a temporary [Nix]{.pkg} shell to build
    the `default.nix` file.
2.  It then executes `gen-pipeline.R` inside the newly defined environment. This
  triggers [Nix]{.pkg} to run the entire polyglot pipeline in the correct order.

Upon completion, the script will have generated all intermediate artefacts and
the final `readme.html` report found in the `pipeline-output` folder containing
the visualisation.